---
title: "01_AM_preprocessing_v1"
author: "Esteban Correa"
date: "12/3/2021"
output: html_document
---

v1b
Age*gender interaction added

-v2
long DF to wide DF, dates for first diagnose of each 5 diseases. Heuristic logic for basic atopic trajectories.

v4
interval type dataset for AS,AR,AD,FA

v5
include addresses per visit, and greenspace

v6
filepath changes to work properly with degauss outputs.
EoE has been added as 5th disease

v7
public insurance added to the dataset

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
memory.size()
rm(list = ls())
memory.limit(size=64000)
gc()
library(tidyverse)
library(foreign)
library(lubridate)
library(gtsummary)
library(survminer)
library(survival)

# diag <- read.table("../../EHR-CCHMC 2009-2020/1426_Visit_diagnosis.txt",sep="|",header = T,colClasses ="character")
# diag$PAT_MRN_ID<-sprintf("%010d", as.numeric(diag$PAT_MRN_ID))
# #
# patients<-read.delim("../../EHR-CCHMC 2009-2020/1426_Demo.txt",header = T,sep = "|")
# patients$PATIENT_MRN<-sprintf("%010d", as.numeric(patients$PATIENT_MRN))
# colnames(patients)[1]<-"PAT_MRN_ID"
# 
# addresses<-read.delim("../../EHR-CCHMC 2009-2020/1426_Address_list.txt",header = T,sep = "|")
# addresses$PAT_MRN_ID<-sprintf("%010d", as.numeric(addresses$PAT_MRN_ID))
# addresses<-addresses %>% 
#   filter(STATE_HX %in% c("Kentucky","Ohio")) %>% 
#   mutate(VISIT_DATE_adr=mdy(VISIT_DATE),
#          id_date=paste0(PAT_MRN_ID,"-",VISIT_DATE_adr)) %>% 
#   distinct(id_date,.keep_all = T) %>% 
#   mutate(fullAddress=paste(ADDR_HX_LINE1,CITY_HX,STATE_HX,ZIP_HX))%>% 
#   dplyr::select(id_date,VISIT_DATE_adr,fullAddress)


# save.image("files.RData")
load("data/files.RData")
```




J30 Vasomotor and allergic rhinitis

J45 Asthma

K20.0 Eosinophilic esophagitis

L20 Atopic dermatitis

Z68.54 greater than or equal to 95th percentile for age

Z77.22 Contact with and (suspected) exposure to environmental tobacco smoke (acute) (chronic)

<!-- Z81.2 Family history of tobacco abuse and dependence -->

Z82.5 family history of asthma and other chronic lower respiratory diseases

Z84.0 Family history of subcutaneous diseases

Z84.89 Family history of allergy

Z91.01 Food allergy status

```{r}
diseases<-'J30|J45|K20.0|L20|Z91.01'

diag2<-diag%>%
  dplyr::filter(grepl(diseases,
                      CURRENT_ICD10_LIST,
                      ignore.case = T) & !grepl('O99|J20|J44',
                      CURRENT_ICD10_LIST,
                      ignore.case = T)) %>%

  mutate(VISIT_DATE2=lubridate::mdy(VISIT_DATE),
         myICD=str_extract(CURRENT_ICD10_LIST,
                             regex("[A-Z][0-9]+\\.[0-9]",
                                   ignore_case = TRUE))) %>% 
  mutate(myICD=ifelse(myICD=='Z84.0',"Z99.0",myICD)) %>% 
  # mutate(myICD=as.factor(recode(myICD,`Z84.0` = "Z99.0"))) %>% 
  mutate(myICD=str_extract(myICD,
                             regex("[A-Z][0-9]+",
                                   ignore_case = TRUE))) %>% 
  mutate(myICD=as.factor(recode(myICD,`E66` = "Z68",`H10`="J30")))
  

summary(as.factor(diag2$myICD))


comorbidities_list<-'Z68.54|Z77.22|Z82.5|Z84.0|Z84.89'

comorb1<-diag%>%
  dplyr::filter(grepl(comorbidities_list,
                      CURRENT_ICD10_LIST,
                      ignore.case = T) & !grepl('O99|J20|J44',
                      CURRENT_ICD10_LIST,
                      ignore.case = T)) %>%

  mutate(VISIT_DATE2=lubridate::mdy(VISIT_DATE),
         myICD=str_extract(CURRENT_ICD10_LIST,
                             regex("[A-Z][0-9]+\\.[0-9]",
                                   ignore_case = TRUE))) %>% 
  mutate(myICD=ifelse(myICD=='Z84.0',"Z99.0",myICD)) %>% 
  # mutate(myICD=as.factor(recode(myICD,`Z84.0` = "Z99.0"))) %>% 
  mutate(myICD=str_extract(myICD,
                             regex("[A-Z][0-9]+",
                                   ignore_case = TRUE))) %>% 
  mutate(myICD=as.factor(recode(myICD,`E66` = "Z68",`H10`="J30")))
  

summary(as.factor(comorb1$myICD))


mindate=ymd('20150101')
maxdate=ymd('20191231') # we should not include covid year
(x=interval((mindate),(maxdate)))
# (x_m= x %/% months(1))
(x_m= 120)
x_days=4383

```

## Limit those visits until 21yo

```{r}
patients2<-patients %>% 
  dplyr::select(PAT_MRN_ID,BIRTH_DATE,GENDER,PATIENT_RACE,ETHNICITY,CURRENT_ADDRESS,CURRENT_CITY,CURRENT_STATE,CURRENT_ZIP) %>% 
  filter(PATIENT_RACE %in% c("WHITE","BLACK OR AFRICAN AMERICAN")) %>% 
  filter(CURRENT_STATE %in% c("Ohio","Kentucky"))
summary(as.factor(patients2$PATIENT_RACE))

diag3<-diag2 %>% 
  left_join(patients2,by="PAT_MRN_ID") %>% 
  drop_na(BIRTH_DATE)

diag3$BIRTH_DATE2<-mdy(diag3$BIRTH_DATE)
# summary(diag3)

diag4<-diag3 %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  drop_na(BIRTH_DATE2)

# summary(diag4)

diag4$age_visit=round((unclass(diag4$VISIT_DATE2)-unclass(diag4$BIRTH_DATE2)) / 365.25,1)

diag5<-diag4 %>% 
  filter(age_visit>=0 & age_visit<21)
summary(diag5)
```

Filter those similar diagnosis that occurred in less than 7 days:

```{r}
# estimate date difference in days
diag6<-diag5%>%
  select(PAT_MRN_ID,VISIT_DATE2,myICD,GENDER,PATIENT_RACE,age_visit,BIRTH_DATE2) %>% 
  dplyr::group_by(PAT_MRN_ID,myICD)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2,myICD) %>% 
  mutate(date_diff2 = as.numeric(VISIT_DATE2)-lag(as.numeric(VISIT_DATE2)))
  # head(.,1000)  # QUICK TEST FOR FIRST TEN ROWS
  
# manual test of dates
# as.numeric(ymd("2015-10-21"))-as.numeric(ymd("2015-07-17"))
# as.numeric(ymd("2015-10-23"))-as.numeric(ymd("2015-07-17"))

# Filter those first and >7 days similar diagnosis
diag7<-diag6 %>% 
  filter(is.na(date_diff2) | date_diff2>7)   %>% 
  filter(VISIT_DATE2>=mindate & VISIT_DATE2<=maxdate)

# all patients previously diagnosed to mindate
prevdiag7<-diag6 %>% 
  filter(is.na(date_diff2) | date_diff2>7)   %>% 
  filter(VISIT_DATE2<mindate) %>% 
  select(PAT_MRN_ID) %>% 
  ungroup() %>% 
  distinct(PAT_MRN_ID)

#exclude from diag7 all patients previously diagnosed to mindate
diag8<-diag7 %>% 
  dplyr::filter(!(PAT_MRN_ID %in% prevdiag7$PAT_MRN_ID))







diag7$interval=round(time_length(interval(mindate,diag7$VISIT_DATE2),"months"),2)# join dates information
diag7$id_date<-paste0(diag7$PAT_MRN_ID,"-",diag7$VISIT_DATE2)


```


```{r}

comorb1<-diag7 %>% 
  # dplyr::filter(myICD!="J45") %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         # prevEoE=if_else(myICD=="K20",1,0),
         # EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="J45" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA)


```


# Degauss

## Degauss: geocode

This part creates the pipelines for geocoding, deprivation, proximity to roads, greenspace and daily PM.

```{r}
degauss<-addresses %>% 
  ungroup() %>% 
  filter(VISIT_DATE_adr>=mindate & VISIT_DATE_adr<=maxdate) %>% 
  mutate(start_date=VISIT_DATE_adr-6) %>% 
  rename(
         end_date=VISIT_DATE_adr,
         address=fullAddress) %>% 
  dplyr::select(id_date,address,start_date,end_date) %>% 
  distinct(id_date,address,.keep_all = T)


degauss2<-degauss %>% 
  right_join(diag7 %>% select(id_date) ,by ="id_date") %>% 
  drop_na(address)

# %>% 
#   dplyr::select(id_date,address,start_date,end_date) 
sum(is.na(degauss2$address))
## address to geocode 
#split
# str(split(degauss, (seq(nrow(degauss))-1) %/% 500) )
files<-(split(degauss2, (seq(nrow(degauss2))-1) %/%150000) )
files
i<-0
for (idx in files){
  idx$start_date<-as.character(idx$start_date)
  idx$end_date<-as.character(idx$end_date)
  write.csv(idx,file = paste0("data/degauss/atmv5_",i,"_base.csv"),row.names = F)
  i<-i+1
}
```

```{}
# please open the terminal, go to the folder (no spaces) and type:
# docker run --rm -v $PWD:/tmp ghcr.io/degauss-org/geocoder:3.2.1 atmv5_2_base.csv 
```

bind all separated geocoded files

```{r}
# find out all geocoded files
geocoded_files<-list.files(path = "data/degauss/", full.names = FALSE,
                     pattern = "geocoder_3.2.1")
geocoded_files
#combine them
degauss3<-{}
for (currentFilename in geocoded_files){
  temp<-read.csv(paste0("data/degauss/",currentFilename))
  degauss3<-rbind(degauss3,temp)
  # print(currentFilename)
}

# filter those correctly geocoded
degauss4<-degauss3 %>% 
  drop_na(lat) %>% 
  dplyr::select(id_date,start_date,end_date,lat,lon)
```

# Degauss: deprivation

```{r}
# str(split(degauss, (seq(nrow(degauss))-1) %/% 500) )
files<-(split(degauss4, (seq(nrow(degauss4))-1) %/%150000) )
i<-0
for (idx in files){
  idx$start_date<-as.character(idx$start_date)
  idx$end_date<-as.character(idx$end_date)
  write.csv(idx,file = paste0("data/degauss/atmv5_",i,"_geocoded.csv"),row.names = F)
  i<-i+1
}
```


```{}
# please open the terminal, go to the folder (no spaces) and type:
docker run --rm -v $PWD:/tmp ghcr.io/degauss-org/dep_index:0.2.1 atmv5_5_geocoded.csv 
```

bind all separated dep_index files

```{r}
# find out all geocoded files
depidx_files<-list.files(path = "data/degauss/", full.names = FALSE,
                     pattern = "dep_index")
depidx_files
#combine them
degauss5<-{}
for (currentFilename in depidx_files){
  temp<-read.csv(paste0("data/degauss/",currentFilename))
  degauss5<-rbind(degauss5,temp)
}

# Filter by
# Butler, Warren, Hamilton, Clermont,
# Boone, kenton, Campbell
coi.list<-c("39017","39165","39061","39025",
            "21015","21117","21037")
degauss6<-degauss5 %>% 
  select(id_date,start_date,end_date,census_tract_id,dep_index,lat,lon) %>% 
  dplyr::mutate(county_id=as.factor(substr(census_tract_id, 1, 5))) %>% 
  dplyr::filter(county_id %in% coi.list) %>% 
  distinct(id_date,.keep_all=T)       # to avoid same ID and date 
# filter those correctly geocoded


```

# degauss: Greenspace

```{r}
degauss7<-degauss6 %>% 
  select(id_date,lat,lon)

write.csv(degauss7,file = paste0("data/degauss/atmv5_all_dep_index.csv"),row.names = F)


# find out all greenspace files
greenspace_files<-list.files(path = "data/degauss/", full.names = FALSE,
                     pattern = "greenspace_0.3.0.csv")
greenspace_files
#combine them
degauss_ndvi<-{}
for (currentFilename in greenspace_files){
  temp<-read.csv(paste0("data/degauss/",currentFilename))
  degauss_ndvi<-rbind(degauss_ndvi,temp)
}

degauss_ndvi<-degauss_ndvi %>% 
  mutate(ndvi=rowMeans(select(., evi_500:evi_2500),na.rm = T)) %>% 
  select(id_date,ndvi)


```

```{}
# please open the terminal, go to the folder (no spaces) and type:
docker run --rm -v $PWD:/tmp ghcr.io/degauss-org/greenspace:0.3.0 atmv5_all_dep_index.csv
```


# degauss: PM

```{r}
#write files
# str(split(degauss, (seq(nrow(degauss))-1) %/% 500) )
files<-(split(degauss6, (seq(nrow(degauss6))-1) %/%140000) )
i<-0
for (idx in files){
  idx$start_date<-as.character(idx$start_date)
  idx$end_date<-as.character(idx$end_date)
  write.csv(idx,file = paste0("data/degauss/atmv5_",i,"_dep_idx.csv"),row.names = F)
  i<-i+1
}

```

```{}
# please open the terminal, go to the folder (no spaces) and type:
docker run --rm -v $PWD:/tmp ghcr.io/degauss-org/pm:0.2.0 atmv5_0_dep_idx.csv
```

Then we take splitted pm files and join into one full averaged pm file:

```{r}

pm_files<-list.files(path = "data/degauss/", full.names = FALSE,
                     pattern = "pm_0.2.0.csv")
pm_files

degauss_pm<-{}
for (currentFilename in pm_files){
  
  temp<-read.csv(paste0("data/degauss/",currentFilename))
  degauss_pm<-rbind(degauss_pm,temp)
  # print(currentFilename)
}

# average daily measures of PM by PAT_MRN_ID and date
degauss_pm2<-degauss_pm%>%
    dplyr::group_by(id_date)%>%
    dplyr::summarize(nsamplesPM=n(),
                     year=first(year),
                     avgDailyPM=mean(pm_pred,na.rm = T))


hist(degauss_pm2$avgDailyPM,64)
```

Create exposures data frame

```{r}
exposures<-degauss6 %>% 
  inner_join(degauss_ndvi,by="id_date") %>% 
  inner_join(degauss_pm2,by="id_date")

# write.csv(exposures,file = paste0("data/exposures_v1.csv"),row.names = F)

exposures<-read.csv("data/exposures_v1.csv")
```

#AD

```{r}
diag8 <- diag7 %>%
  dplyr::filter(myICD == "L20") %>%
  dplyr::group_by(PAT_MRN_ID, myICD) %>%
  arrange(PAT_MRN_ID, myICD, VISIT_DATE2) %>%
  mutate(futime = 60, visitNum = paste0("visit", sprintf("%03d", row_number())))
diag8_wide <-
  spread(diag8 %>% select(PAT_MRN_ID, interval, visitNum),
         visitNum,
         interval) 

diag8_wide$futime=60
visits<-diag8_wide %>% select(starts_with("visit"))
(visits<-names(visits[,-c(1,2)]))

diag9_wide<-tmerge(diag8_wide %>% select(-starts_with("visit")), diag8_wide, id=PAT_MRN_ID, tstop=futime)

for(name in visits){
  # print(name)
  diag9_wide <- tmerge(diag9_wide, diag8_wide, id=PAT_MRN_ID, edvisit = event(get(name))
                     )
}
#55632724
diag9_wide <- tmerge(diag9_wide, diag9_wide, id=PAT_MRN_ID, enum=cumtdc(tstart))
diag9_wide$id_interval=paste0(diag9_wide$PAT_MRN_ID,"_",sprintf("%.2f", diag9_wide$tstop))
```

Joining

```{r}
# join comorbidities
comorb1<-diag7 %>% 
  # dplyr::filter(myICD!="J45") %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         prevEoE=if_else(myICD=="K20",1,0),
         EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="L20" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA,EoE)

diag10_wide<-diag9_wide %>% 
  left_join(comorb1,by="id_interval")

# Join VISIT_DATE2 by id_interval
diag8_visitdate<-diag8 %>% 
  ungroup() %>% 
  select(PAT_MRN_ID,interval,VISIT_DATE2,id_date) %>% 
  mutate(id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))) %>% 
  select(VISIT_DATE2,id_interval,id_date)

diag11_wide<-diag10_wide %>% 
  left_join(diag8_visitdate,by="id_interval") 


#Join of exposures
diag12_wide<-diag11_wide %>% 
  left_join(exposures,by="id_date") 

# NA filling
diag13_wide <- diag12_wide %>%
  dplyr::group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID, tstart) %>%
  mutate(
    BIRTH_DATE2 = if_else(is.na(BIRTH_DATE2), lag(BIRTH_DATE2), BIRTH_DATE2),
    VISIT_DATE2 = if_else(is.na(VISIT_DATE2), maxdate, VISIT_DATE2),
    RACE = if_else(is.na(PATIENT_RACE), lag(PATIENT_RACE), PATIENT_RACE),
    GENDER = if_else(is.na(GENDER), lag(GENDER), GENDER),
    AD = if_else(is.na(AD), lag(AD), AD),
    AS = if_else(is.na(AS), lag(AS), AS),
    AR = if_else(is.na(AR), lag(AR), AR),
    FA = if_else(is.na(FA), lag(FA), FA),
    EoE = if_else(is.na(EoE), lag(EoE), EoE),
    age_visit=if_else(is.na(age_visit),round((unclass(maxdate)-unclass(BIRTH_DATE2)) / 365.25,1),age_visit),
    lat=if_else(is.na(lat), lag(lat), lat) # used as a flag for deletion
  ) %>%  
  mutate(RACE=as.factor(recode(RACE, 
                         `WHITE`="EA",
                         `BLACK OR AFRICAN AMERICAN`="AA"))) 

# test with this case 0000637772

ad_wide<-diag13_wide%>% 
  select(PAT_MRN_ID,VISIT_DATE2,id_date,id_interval,tstart,tstop,edvisit,futime,enum,age_visit,GENDER,RACE,AD,AS,AR,FA,EoE,
         lat,lon,census_tract_id,county_id,avgDailyPM,dep_index,ndvi) %>% 
  mutate(year=year(VISIT_DATE2)) %>% 
  filter(!is.na(lat) )
# write.csv(ad_wide,file = paste0("data/ad_wide_v1.csv"),row.names = F)
```



# Asthma

```{r}
# asthma
diag8<-diag7 %>% 
  dplyr::filter(myICD=="J45") %>% 
  dplyr::group_by(PAT_MRN_ID,myICD)%>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>% 
  mutate(futime=60,visitNum=paste0("visit",sprintf("%03d", row_number())))
  

diag8_wide <- spread(diag8%>% select(PAT_MRN_ID,interval,visitNum), visitNum, interval) 

diag8_wide$futime=60
visits<-diag8_wide %>% select(starts_with("visit"))
(visits<-names(visits[,-c(1,2)]))

diag9_wide<-tmerge(diag8_wide %>% select(-starts_with("visit")), diag8_wide, id=PAT_MRN_ID, tstop=futime)

for(name in visits){
  # print(name)
  diag9_wide <- tmerge(diag9_wide, diag8_wide, id=PAT_MRN_ID, edvisit = event(get(name))
                     )
}
#55632724
diag9_wide <- tmerge(diag9_wide, diag9_wide, id=PAT_MRN_ID, enum=cumtdc(tstart))
diag9_wide$id_interval=paste0(diag9_wide$PAT_MRN_ID,"_",sprintf("%.2f", diag9_wide$tstop))
```

Joining

```{r}
# join comorbidities
comorb1<-diag7 %>% 
  # dplyr::filter(myICD!="J45") %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         prevEoE=if_else(myICD=="K20",1,0),
         EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="J45" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA,EoE)

diag10_wide<-diag9_wide %>% 
  left_join(comorb1,by="id_interval")

# Join VISIT_DATE2 by id_interval
diag8_visitdate<-diag8 %>% 
  select(PAT_MRN_ID,interval,VISIT_DATE2,id_date) %>% 
  mutate(id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval)))

diag11_wide<-diag10_wide %>% 
  left_join(diag8_visitdate,by="id_interval") %>% 
  select(-c(PAT_MRN_ID.y,interval.y,myICD.y)) %>% 
  rename(
         myICD=myICD.x,
         PAT_MRN_ID=PAT_MRN_ID.x,
         interval=interval.x) 

#Join of exposures
diag12_wide<-diag11_wide %>% 
  left_join(exposures,by="id_date") 

# NA filling
diag13_wide <- diag12_wide %>%
  dplyr::group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID, tstart) %>%
  mutate(
    BIRTH_DATE2 = if_else(is.na(BIRTH_DATE2), lag(BIRTH_DATE2), BIRTH_DATE2),
    VISIT_DATE2 = if_else(is.na(VISIT_DATE2), maxdate, VISIT_DATE2),
    RACE = if_else(is.na(PATIENT_RACE), lag(PATIENT_RACE), PATIENT_RACE),
    GENDER = if_else(is.na(GENDER), lag(GENDER), GENDER),
    AD = if_else(is.na(AD), lag(AD), AD),
    AS = if_else(is.na(AS), lag(AS), AS),
    AR = if_else(is.na(AR), lag(AR), AR),
    FA = if_else(is.na(FA), lag(FA), FA),
    EoE = if_else(is.na(EoE), lag(EoE), EoE),
    age_visit=if_else(is.na(age_visit),round((unclass(maxdate)-unclass(BIRTH_DATE2)) / 365.25,1),age_visit),
    lat=if_else(is.na(lat), lag(lat), lat) # used as a flag for deletion
  ) %>%  
  mutate(RACE=as.factor(recode(RACE, 
                         `WHITE`="EA",
                         `BLACK OR AFRICAN AMERICAN`="AA"))) 

# test with this case 0000637772

as_wide<-diag13_wide%>% 
  select(PAT_MRN_ID,VISIT_DATE2,id_date,id_interval,tstart,tstop,edvisit,futime,enum,age_visit,GENDER,RACE,AD,AS,AR,FA,EoE,
         lat,lon,census_tract_id,county_id,avgDailyPM,dep_index,ndvi) %>% 
  mutate(year=year(VISIT_DATE2)) %>% 
  filter(!is.na(lat) )

# write.csv(as_wide,file = paste0("data/as_wide_v1.csv"),row.names = F)
```

# EoE

```{r}
# asthma
diag8<-diag7 %>% 
  dplyr::filter(myICD=="K20") %>% 
  dplyr::group_by(PAT_MRN_ID,myICD)%>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>% 
  mutate(futime=60,visitNum=paste0("visit",sprintf("%03d", row_number())))
  

diag8_wide <- spread(diag8%>% select(PAT_MRN_ID,interval,visitNum), visitNum, interval) 

diag8_wide$futime=60
visits<-diag8_wide %>% select(starts_with("visit"))
(visits<-names(visits[,-c(1,2)]))

diag9_wide<-tmerge(diag8_wide %>% select(-starts_with("visit")), diag8_wide, id=PAT_MRN_ID, tstop=futime)

for(name in visits){
  # print(name)
  diag9_wide <- tmerge(diag9_wide, diag8_wide, id=PAT_MRN_ID, edvisit = event(get(name))
                     )
}
#55632724
diag9_wide <- tmerge(diag9_wide, diag9_wide, id=PAT_MRN_ID, enum=cumtdc(tstart))
diag9_wide$id_interval=paste0(diag9_wide$PAT_MRN_ID,"_",sprintf("%.2f", diag9_wide$tstop))
```

Joining all to EoE

```{r}
# join comorbidities
comorb1<-diag7 %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         prevEoE=if_else(myICD=="K20",1,0),
         EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="K20" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA,EoE)

diag10_wide<-diag9_wide %>% 
  left_join(comorb1,by="id_interval")

# Join VISIT_DATE2 by id_interval
diag8_visitdate<-diag8 %>% 
  select(PAT_MRN_ID,interval,VISIT_DATE2,id_date) %>% 
  mutate(id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval)))

diag11_wide<-diag10_wide %>% 
  left_join(diag8_visitdate,by="id_interval") %>% 
  select(-c(PAT_MRN_ID.y,interval.y,myICD.y)) %>% 
  rename(
         myICD=myICD.x,
         PAT_MRN_ID=PAT_MRN_ID.x,
         interval=interval.x) 

#Join of exposures
diag12_wide<-diag11_wide %>% 
  left_join(exposures,by="id_date") 

# NA filling
diag13_wide <- diag12_wide %>%
  dplyr::group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID, tstart) %>%
  mutate(
    BIRTH_DATE2 = if_else(is.na(BIRTH_DATE2), lag(BIRTH_DATE2), BIRTH_DATE2),
    VISIT_DATE2 = if_else(is.na(VISIT_DATE2), maxdate, VISIT_DATE2),
    RACE = if_else(is.na(PATIENT_RACE), lag(PATIENT_RACE), PATIENT_RACE),
    GENDER = if_else(is.na(GENDER), lag(GENDER), GENDER),
    AD = if_else(is.na(AD), lag(AD), AD),
    AS = if_else(is.na(AS), lag(AS), AS),
    AR = if_else(is.na(AR), lag(AR), AR),
    FA = if_else(is.na(FA), lag(FA), FA),
    EoE = if_else(is.na(EoE), lag(EoE), EoE),
    age_visit=if_else(is.na(age_visit),round((unclass(maxdate)-unclass(BIRTH_DATE2)) / 365.25,1),age_visit),
    lat=if_else(is.na(lat), lag(lat), lat) # used as a flag for deletion
  ) %>%  
  mutate(RACE=as.factor(recode(RACE, 
                         `WHITE`="EA",
                         `BLACK OR AFRICAN AMERICAN`="AA"))) 

# test with this case 0000637772

eoe_wide<-diag13_wide%>% 
  select(PAT_MRN_ID,VISIT_DATE2,id_date,id_interval,tstart,tstop,edvisit,futime,enum,age_visit,GENDER,RACE,AD,AS,AR,FA,EoE,
         lat,lon,census_tract_id,county_id,avgDailyPM,dep_index,ndvi) %>% 
  mutate(year=year(VISIT_DATE2)) %>% 
  filter(!is.na(lat) )

# write.csv(eoe_wide,file = paste0("data/as_wide_v1.csv"),row.names = F)
```


```{r}
library(tableone)
# (table(diag12$visit_1,diag12$RACE))
diag12_AS<-as_wide
diag12_AS<-droplevels(diag12_AS)


myVars<-c("dep_index","avgDailyPM")
tab <- CreateTableOne(vars = myVars, 
                      # strata = "RACE" ,
                      data = diag12_AS,
                       includeNA = T,
                          test = T)
# tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
# tab2Mat
kableone(tab,nonnormal=T)
tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = diag12_AS,
                       includeNA = T,
                          test = T)
# tab2Mat<-print(tab showAllLevels = T, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
# tab2Mat
kableone(tab,nonnormal=T)
```


## Depidx at visit

```{r}
var="dep_index"

ggplot()+
  geom_point(data=as_wide,aes(x=year,y=get(var),col=RACE),position = "jitter",alpha=0.05,size=0.1)+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean,
               geom="line",size=1,position=position_dodge(0.2))+
  scale_x_continuous(breaks = c(2015,2016,2017,2018,2019))+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean, geom="point",
               size=1.5,position=position_dodge(0.2))+
  labs(subtitle = "",x="",y=var,color="Race")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),legend.position = "bottom")
```

## ndvi at visit

```{r}
var="ndvi"

ggplot()+
  geom_point(data=as_wide,aes(x=year,y=get(var),col=RACE),position = "jitter",alpha=0.05,size=0.1)+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean,
               geom="line",size=1,position=position_dodge(0.2))+
  scale_x_continuous(breaks = c(2015,2016,2017,2018,2019))+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean, geom="point",
               size=1.5,position=position_dodge(0.2))+
  labs(subtitle = "",x="",y=var,color="Race")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),legend.position = "bottom")

```


## PM at visit

```{r}
var="avgDailyPM"

ggplot()+
  geom_point(data=as_wide,aes(x=year,y=get(var),col=RACE),position = "jitter",alpha=0.05,size=0.1)+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean,
               geom="line",size=1,position=position_dodge(0.2))+
  scale_x_continuous(breaks = c(2015,2016,2017,2018,2019))+
  stat_summary(data=as_wide,aes(x=year,y=get(var),col=RACE), fun.y=mean, geom="point",
               size=1.5,position=position_dodge(0.2))+
  labs(subtitle = "",x="",y=var,color="Race")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),legend.position = "bottom")

```

```{r}
library(survival)



data<-as_wide %>% 
  # dplyr::filter(myICD=="J45") %>% 
  mutate(RACE=factor(recode(RACE,`WHITE`="EA",`BLACK OR AFRICAN AMERICAN`="AA"),levels=c("EA","AA"))
         ) %>% drop_na(age_visit)
summary(data)

data %>% 
  distinct(PAT_MRN_ID,.keep_all = T) %>% 
  group_by(RACE) %>% 
  summarise(freq=n())

f1=formula(Surv(tstart, tstop, edvisit) ~ RACE+AD+FA+AR+age_visit*GENDER)
cox.as<-coxph(formula = f1,data =data)
summary(cox.as)

f1=formula(Surv(tstart, tstop, edvisit) ~ RACE+AD+age_visit*GENDER+dep_index+avgDailyPM+ndvi)
cox.as<-coxph(formula = f1,data =data)

summary(cox.as)

```

# Allergic rhinitis

```{r}
# AR
diag8<-diag7 %>% 
  dplyr::filter(myICD=="J30") %>% 
  dplyr::group_by(PAT_MRN_ID,myICD)%>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>% 
  mutate(futime=60,visitNum=paste0("visit",sprintf("%03d", row_number())))
  

diag8_wide <- spread(diag8%>% select(PAT_MRN_ID,interval,visitNum), visitNum, interval) 

diag8_wide$futime=60
visits<-diag8_wide %>% select(starts_with("visit"))
(visits<-names(visits[,-c(1,2)]))

diag9_wide<-tmerge(diag8_wide %>% select(-starts_with("visit")), diag8_wide, id=PAT_MRN_ID, tstop=futime)

for(name in visits){
  # print(name)
  diag9_wide <- tmerge(diag9_wide, diag8_wide, id=PAT_MRN_ID, edvisit = event(get(name))
                     )
}
#55632724
diag9_wide <- tmerge(diag9_wide, diag9_wide, id=PAT_MRN_ID, enum=cumtdc(tstart))
diag9_wide$id_interval=paste0(diag9_wide$PAT_MRN_ID,"_",sprintf("%.2f", diag9_wide$tstop))
```

Joining

```{r}
# join comorbidities
comorb1<-diag7 %>% 
  # dplyr::filter(myICD!="J45") %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         prevEoE=if_else(myICD=="K20",1,0),
         EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="J30" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA,EoE)

diag10_wide<-diag9_wide %>% 
  left_join(comorb1,by="id_interval")

# Join VISIT_DATE2 by id_interval
diag8_visitdate<-diag8 %>% 
  select(PAT_MRN_ID,interval,VISIT_DATE2,id_date) %>% 
  mutate(id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval)))

diag11_wide<-diag10_wide %>% 
  left_join(diag8_visitdate,by="id_interval") %>% 
  select(-c(PAT_MRN_ID.y,interval.y,myICD.y)) %>% 
  rename(
         myICD=myICD.x,
         PAT_MRN_ID=PAT_MRN_ID.x,
         interval=interval.x) 

#Join of exposures
diag12_wide<-diag11_wide %>% 
  left_join(exposures,by="id_date") 

# NA filling
diag13_wide <- diag12_wide %>%
  dplyr::group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID, tstart) %>%
  mutate(
    BIRTH_DATE2 = if_else(is.na(BIRTH_DATE2), lag(BIRTH_DATE2), BIRTH_DATE2),
    VISIT_DATE2 = if_else(is.na(VISIT_DATE2), maxdate, VISIT_DATE2),
    RACE = if_else(is.na(PATIENT_RACE), lag(PATIENT_RACE), PATIENT_RACE),
    GENDER = if_else(is.na(GENDER), lag(GENDER), GENDER),
    AD = if_else(is.na(AD), lag(AD), AD),
    AS = if_else(is.na(AS), lag(AS), AS),
    AR = if_else(is.na(AR), lag(AR), AR),
    FA = if_else(is.na(FA), lag(FA), FA),
    EoE = if_else(is.na(EoE), lag(EoE), EoE),
    age_visit=if_else(is.na(age_visit),round((unclass(maxdate)-unclass(BIRTH_DATE2)) / 365.25,1),age_visit),
    lat=if_else(is.na(lat), lag(lat), lat) # used as a flag for deletion
  ) %>%  
  mutate(RACE=as.factor(recode(RACE, 
                         `WHITE`="EA",
                         `BLACK OR AFRICAN AMERICAN`="AA"))) 

# test with this case 0000637772

ar_wide<-diag13_wide%>% 
  select(PAT_MRN_ID,VISIT_DATE2,id_date,id_interval,tstart,tstop,edvisit,futime,enum,age_visit,GENDER,RACE,AD,AS,AR,FA,EoE,
         lat,lon,census_tract_id,county_id,avgDailyPM,dep_index,ndvi) %>% 
  mutate(year=year(VISIT_DATE2)) %>% 
  filter(!is.na(lat) )

```


# FA

```{r}
# FA
diag8<-diag7 %>% 
  dplyr::filter(myICD=="Z91") %>% 
  dplyr::group_by(PAT_MRN_ID,myICD)%>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>% 
  mutate(futime=60,visitNum=paste0("visit",sprintf("%03d", row_number())))
  
diag8_wide <- spread(diag8%>% select(PAT_MRN_ID,interval,visitNum), visitNum, interval) 

diag8_wide$futime=60
visits<-diag8_wide %>% select(starts_with("visit"))
(visits<-names(visits[,-c(1,2)]))

diag9_wide<-tmerge(diag8_wide %>% select(-starts_with("visit")), diag8_wide, id=PAT_MRN_ID, tstop=futime)

for(name in visits){
  # print(name)
  diag9_wide <- tmerge(diag9_wide, diag8_wide, id=PAT_MRN_ID, edvisit = event(get(name))
                     )
}
#55632724
diag9_wide <- tmerge(diag9_wide, diag9_wide, id=PAT_MRN_ID, enum=cumtdc(tstart))
diag9_wide$id_interval=paste0(diag9_wide$PAT_MRN_ID,"_",sprintf("%.2f", diag9_wide$tstop))
```


## Joining FA to comorbities

```{r}
# join comorbidities
comorb1<-diag7 %>% 
  dplyr::group_by(PAT_MRN_ID)%>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>%
  # head(.,1000) %>%   # QUICK TEST FOR FIRST TEN ROWS
  mutate(prevAD=if_else(myICD=="L20",1,0),
         AD=if_else(cumsum(prevAD)>0,T,F),
         prevAS=if_else(myICD=="J45",1,0),
         AS=if_else(cumsum(prevAS)>0,T,F),
         prevAR=if_else(myICD=="J30",1,0),
         AR=if_else(cumsum(prevAR)>0,T,F),
         prevFA=if_else(myICD=="Z91",1,0),
         FA=if_else(cumsum(prevFA)>0,T,F),
         prevEoE=if_else(myICD=="K20",1,0),
         EoE=if_else(cumsum(prevEoE)>0,T,F),
         id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval))
         ) %>% 
  dplyr::filter(myICD=="Z91" & interval>=0) %>% 
  ungroup() %>% 
  select(id_interval,BIRTH_DATE2,GENDER,PATIENT_RACE,age_visit,interval,AD,AS,AR,FA,EoE)

diag10_wide<-diag9_wide %>% 
  left_join(comorb1,by="id_interval")

# Join VISIT_DATE2 by id_interval
diag8_visitdate<-diag8 %>% 
  select(PAT_MRN_ID,interval,VISIT_DATE2,id_date) %>% 
  mutate(id_interval=paste0(PAT_MRN_ID,"_",sprintf("%.2f", interval)))

diag11_wide<-diag10_wide %>% 
  left_join(diag8_visitdate,by="id_interval") %>% 
  select(-c(PAT_MRN_ID.y,interval.y,myICD.y)) %>% 
  rename(
         myICD=myICD.x,
         PAT_MRN_ID=PAT_MRN_ID.x,
         interval=interval.x) 

#Join of exposures
diag12_wide<-diag11_wide %>% 
  left_join(exposures,by="id_date") 

# NA filling
diag13_wide <- diag12_wide %>%
  dplyr::group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID, tstart) %>%
  mutate(
    BIRTH_DATE2 = if_else(is.na(BIRTH_DATE2), lag(BIRTH_DATE2), BIRTH_DATE2),
    VISIT_DATE2 = if_else(is.na(VISIT_DATE2), maxdate, VISIT_DATE2),
    RACE = if_else(is.na(PATIENT_RACE), lag(PATIENT_RACE), PATIENT_RACE),
    GENDER = if_else(is.na(GENDER), lag(GENDER), GENDER),
    AD = if_else(is.na(AD), lag(AD), AD),
    AS = if_else(is.na(AS), lag(AS), AS),
    AR = if_else(is.na(AR), lag(AR), AR),
    FA = if_else(is.na(FA), lag(FA), FA),
    EoE = if_else(is.na(EoE), lag(EoE), EoE),
    age_visit=if_else(is.na(age_visit),round((unclass(maxdate)-unclass(BIRTH_DATE2)) / 365.25,1),age_visit),
    lat=if_else(is.na(lat), lag(lat), lat) # used as a flag for deletion
  ) %>%  
  mutate(RACE=as.factor(recode(RACE, 
                         `WHITE`="EA",
                         `BLACK OR AFRICAN AMERICAN`="AA"))) 

# test with this case 0000637772

fa_wide<-diag13_wide%>% 
  select(PAT_MRN_ID,VISIT_DATE2,id_date,id_interval,tstart,tstop,edvisit,futime,enum,age_visit,GENDER,RACE,AD,AS,AR,FA,EoE,
         lat,lon,census_tract_id,county_id,avgDailyPM,dep_index,ndvi) %>% 
  mutate(year=year(VISIT_DATE2)) %>% 
  filter(!is.na(lat) )

```


```{r}
rm(diag,diag2,diag3,diag4,diag5,diag6,diag7,diag8,diag8_visitdate,diag8_wide,diag9_wide,diag10_wide,
   diag11_wide,diag12_wide,diag13_wide,degauss_ndvi,degauss_pm,degauss_pm2,degauss3,degauss4,
   futime,intervaltime,visits,name,addresses,exposures,
   patients,patients2,diseases,mindate,maxdate,x,x_days,x_m,comorb1,comorbidities_list,
   coi.list,data,temp,degauss5,degauss6,degauss7,cox.as,currentFilename,var,pm_files,f1,depidx_files,greenspace_files)

save.image("data/atm_survival_v6.RData")

```

Adding public insurance

```{r}
hcType<-read.delim("data/1426_Visit_financialclss.txt",header = T,sep = "|")
hcType$PAT_MRN_ID2<-as.integer(hcType$PAT_MRN_ID)

hcType$PAT_MRN_ID3<-sprintf("%010d", hcType$PAT_MRN_ID2)

load("data/atm_survival_v6.RData")
```

```{r}

summary(hcType$FINANCIAL_CLASS)

hcType$hcType<-as.factor(plyr::revalue(as.character(hcType$FINANCIAL_CLASS), 
                                        c("Medicare"="Medicaid",
                                          "Medicaid"="Medicaid",
                                          "Specialty Medicaid"="Medicaid",
                                          "Specialty Medicaid Cap"="Medicaid",
                                          "HMO Medicaid Cap"="Medicaid",
                                          "HMO Medicaid"="Medicaid",
                                          "MACSIS"="Medicaid",
                                          "BCMH"="Medicaid",
                                          "HMO"="Commercial",
                                          "Commercial"="Commercial",
                                          "Blue Shield"="Commercial",
                                          "Exchange"="Commercial",
                                          "CLEARING"="Other",
                                          "International"="Other",
                                          "Self-Pay"="Commercial",
                                          "Specialty"="Commercial",
                                          "Pending Medicaid"="Medicaid"
                                          )))

summary(hcType$hcType)


hc2<-hcType[!duplicated(hcType$PAT_MRN_ID3),]%>%
  dplyr::filter(hcType=="Commercial" | hcType=="Medicaid") %>% 
  select(PAT_MRN_ID3,hcType) %>% 
  rename(PAT_MRN_ID=PAT_MRN_ID3)
hc2<-droplevels(hc2)
summary(hc2$hcType)


```


```{r}
eoe_wide<-left_join(eoe_wide,hc2,by="PAT_MRN_ID")
ad_wide<-left_join(ad_wide,hc2,by="PAT_MRN_ID")
as_wide<-left_join(as_wide,hc2,by="PAT_MRN_ID")
ar_wide<-left_join(ar_wide,hc2,by="PAT_MRN_ID")
fa_wide<-left_join(fa_wide,hc2,by="PAT_MRN_ID")
summary(eoe_wide)
summary(ad_wide)
summary(as_wide)
summary(ar_wide)
summary(fa_wide)
```

```{r}
rm(hc2,hcType,geocoded_files)

save.image("data/atm_survival_v7.RData")

```

