---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


-v1
Using old dataset

-v2
Using preprocesssing_v3 dataset

-v3 
Using preprocesssing_v5 dataset, digits set 1 for age at visit

-v4
use of dataset v6 with EoE

v5
data/atm_survival_v7.RData with public insurance added
jaccard index calculation

v7
drop of deprivation index

v8
non normal t-test

v9
Use of atm_survival_v8.RData with improved inclusion criteria
t-test reemplazados por lmm para time-varying covariates
code clean-up

v10
multimorbidity added 
proportion matrix functions added to helpers.r

v11
sex added code cleanup for reviewer comments
counts for reencounters

v12 
proofread corrections from Dr. Lili Ding:
allergic comorbidities counts added by disease-specific for Table 1

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning = F,results = F)
rm(list = ls())
```



```{r}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(patchwork)
library(tidyverse)
library(foreign)
library(lubridate)
library(tableone)
library(nlme)
source("helpers.r")
cbPalette <- c("#999999",  "#E64B35", "#4DBBD5", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

load("data/atm_survival_v8.RData")
ad_wide$myICD<-"L20"
as_wide$myICD<-"J45"
ar_wide$myICD<-"J30"
fa_wide$myICD<-"Z91"
eoe_wide$myICD<-"K20"
all_diseases<-rbind(ad_wide,as_wide,ar_wide,fa_wide,eoe_wide)%>% 
  mutate(
    comorbidities=rowSums(across(c(AD,AS,AR,FA,EoE))),
    onemore=if_else(comorbidities>=2,"Yes","No"),
    multimorbidity=case_when(
      comorbidities==1 ~ "Zero",
      comorbidities==2 ~ "One",
      comorbidities>=3 ~ "Two_or_more"
      )
  ) %>% 
  filter(edvisit==1)

#799820
#213807
# 69895 censoring
# 283702 all
# write.csv(all_diseases,"data/JACIiP_EHR_Data.csv",row.names = F)
```


```{r}
#first visit
firstVisit<- all_diseases %>%
  filter(enum==1)

#more than one visit
reVisits<- all_diseases %>%
  filter(enum>1 & edvisit!=0)

#first visit
censoredVisit<- all_diseases %>%
  filter(edvisit==0)



```


# Patient-level

```{r}
all_diseases %>% 
  group_by(PAT_MRN_ID) %>% 
  distinct(PAT_MRN_ID,.keep_all = T) %>% 
  summarise(n=n())

all_diseases %>% 
  distinct(PAT_MRN_ID,.keep_all = T) %>% 
  group_by(RACE) %>% 
  summarise(n=n())


all_diseases %>% 
  ungroup() %>% 
  summarise(n=n())

all_diseases %>% 
  group_by(RACE) %>% 
  summarise(n=n())
```



# Sex by disease

```{r}
onlypatients<-all_diseases %>% 
  group_by(PAT_MRN_ID) %>% 
  distinct(PAT_MRN_ID,.keep_all = T)
myVars<-c("GENDER"
          )
tab <- CreateTableOne(
  vars = myVars,
  # strata = "RACE" ,
  data = onlypatients,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T)

tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = onlypatients,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T)
```

```{r}
myVars <- c("multimorbidity")
myVars<-c("AD","FA","AS","AR","EoE")
# CreateContTable is required for limiting digits
data<-all_diseases 
#%>% 
 # filter(myICD=="L20")
tab <- CreateTableOne(
  vars = myVars,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
# kableone(tab,digits=1,nonnormal=T)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
# kableone(tab,digits=1,nonnormal=T)
```


# Age of onset

## AD onset

```{r}
#first diagnosis of each disease
all_onset<-all_diseases %>%
  group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>%
  distinct(myICD,.keep_all = TRUE)

myVars <- c("age_visit")
# CreateContTable is required for limiting digits
tab <- CreateContTable(
  vars = myVars,
  # strata = "RACE" ,
  data = all_onset %>% filter(myICD=="L20"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    digits = 1,
    nonnormal = F,
    showAllLevels = T,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
tab <- CreateContTable(
  vars = myVars,
  strata = "RACE" ,
  data = all_onset %>% filter(myICD=="L20"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    digits=1,
    nonnormal = F,
    showAllLevels = F,
    formatOptions = list(big.mark = ",",digits=1),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
```

## FA onset

```{r}
tab <- CreateContTable(
  vars = myVars,
  # strata = "RACE" ,
  data = all_onset %>% filter(myICD=="Z91"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
tab <- CreateContTable(
  vars = myVars,
  strata = "RACE" ,
  data = all_onset %>% filter(myICD=="Z91"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
```

## AS onset

```{r}
tab <- CreateContTable(
  vars = myVars,
  # strata = "RACE" ,
  data = all_onset %>% filter(myICD=="J45"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
tab <- CreateContTable(
  vars = myVars,
  strata = "RACE" ,
  data = all_onset %>% filter(myICD=="J45"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
```

## AR onset

```{r}
tab <- CreateContTable(
  vars = myVars,
  # strata = "RACE" ,
  data = all_onset %>% filter(myICD=="J30"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
tab <- CreateContTable(
  vars = myVars,
  strata = "RACE" ,
  data = all_onset %>% filter(myICD=="J30"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
```


## EoE onset

```{r}
tab <- CreateContTable(
  vars = myVars,
  # strata = "RACE" ,
  data = all_onset %>% filter(myICD=="K20"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
tab <- CreateContTable(
  vars = myVars,
  strata = "RACE" ,
  data = all_onset %>% filter(myICD=="K20"),
  # includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal = T)
```

# Neighborhood-level


## All diseases at once

```{r}
data<-all_diseases %>% 
  drop_na(ndvi,avgDailyPM,dep_index)
myVars<-c(
          "ndvi","avgDailyPM", "dep_index"
          )
tab <- CreateTableOne(
  vars = myVars,
  # strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T)
```



## AD exposures


```{r}
data<-all_diseases %>% 
  filter(myICD=="L20") %>% 
  drop_na(ndvi,avgDailyPM)
myVars<-c("GENDER","hcType",
          "onemore" # lili's change
          )
tab <- CreateTableOne(
  vars = myVars,
  # strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)


mod_ov <- lme(ndvi ~ 1.0,random=~1|PAT_MRN_ID, data=data)
summary(mod_ov)

mod_race2 <- lme(ndvi ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data)
summary(mod_race2,digits=2)

#SE for overall population 
mod_ov <- lme(avgDailyPM ~ 1.0,random=~1|PAT_MRN_ID, data=data)
summary(mod_ov)

#SE for AA would be the SE for intercept, SE=0.03
mod_race2 <- lme(avgDailyPM ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data)
summary(mod_race2)

#Test of flipping group of reference
#SE for EA would be the intercept, SE=0.04, p-value is the same as AA
data$RACE2<-factor(data$RACE,levels = c("EA","AA"))
mod_race3 <- lme(avgDailyPM ~ 1.0+RACE2,random=~1|PAT_MRN_ID, data=data)
summary(mod_race3)

# no intercept, direct way of getting SE for all groups, no P-value
mod_race3b <- lme(avgDailyPM ~ 0.0+RACE2,random=~1|PAT_MRN_ID, data=data)
summary(mod_race3b)
```


## FA exposures

```{r}
data<-all_diseases %>% 
  filter(myICD=="Z91") 
tab <- CreateTableOne(
  vars = myVars,
  # strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)

mod_ov <- lme(ndvi ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_ov)

mod_race2 <- lme(ndvi ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_race2)
round(0.4248205+0.0285757,2)
mod_ov <- lme(avgDailyPM ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_ov)

mod_race2 <- lme(avgDailyPM ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_race2)

round(8.948805-0.063526,2)
```

## Asthma exposures

```{r}
data<-all_diseases %>% 
  filter(myICD=="J45") 
tab <- CreateTableOne(
  vars = myVars,
  data =data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)

mod_ov <- lme(ndvi ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_ov)
mod_race2 <- lme(ndvi ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_race2)
round(0.4207425+0.0309923,2)

mod_ov <- lme(avgDailyPM ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_ov)
mod_race2 <- lme(avgDailyPM ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_race2)
round(8.850368+0.060346,2)
```

## AR exposures

```{r}
data<-all_diseases %>% 
  filter(myICD=="J30") 
tab <- CreateTableOne(
  vars = myVars,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE,)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE,)

mod_ov <- lme(ndvi ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_ov)
mod_race2 <- lme(ndvi ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_race2)
round(0.4191348+0.0335666,2)

mod_ov <- lme(avgDailyPM ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_ov)
mod_race2 <- lme(avgDailyPM ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_race2)
round(8.730767+0.050879,2)
```

## EoE exposures

```{r}
data<-all_diseases %>% 
  filter(myICD=="K20") 
tab <- CreateTableOne(
  vars = myVars,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)
tab <- CreateTableOne(
  vars = myVars,
  strata = "RACE" ,
  data = data,
  includeNA = T,
  test = T
)
tab2Mat <-
  print(
    tab,
    nonnormal = T,
    showAllLevels = TRUE,
    formatOptions = list(big.mark = ","),
    quote = FALSE,
    noSpaces = TRUE,
    printToggle = FALSE
  )
tab2Mat
kableone(tab,digits=1,nonnormal=T,showAllLevels = TRUE)

mod_ov <- lme(ndvi ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_ov)
mod_race2 <- lme(ndvi ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(ndvi))
summary(mod_race2)
round(0.4313152+0.0284828,2)

mod_ov <- lme(avgDailyPM ~ 1.0,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_ov)
mod_race2 <- lme(avgDailyPM ~ 1.0+RACE,random=~1|PAT_MRN_ID, data=data%>% drop_na(avgDailyPM))
summary(mod_race2)
round(9.086473-0.211349,2)
```

  
## Censoring  

Censoring characterization for RMST

### AD

```{r}
months <-0:59
ad_total<-ad_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(total=n())
t(ad_total)

temp_ad<-ad_total
for (i in months){
  cens<-ad_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  filter(tstart>=i) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise("n.{i}":=n())
  
  temp_ad<-temp_ad %>%
    left_join(cens,by="RACE")
}
temp_ad
temp_ad2<-as.matrix(t(temp_ad))
temp_ad3<-data.frame(month=months,
                     AA=as.numeric(temp_ad2[3:nrow(temp_ad2),1]),
                     EA=as.numeric(temp_ad2[3:nrow(temp_ad2),2])
                     )
temp_ad3$AA_perc<-1-round(temp_ad3$AA/ad_total$total[1],2)
temp_ad3$EA_perc<-1-round(temp_ad3$EA/ad_total$total[2],2)
temp_ad4<-gather(temp_ad3, race, perc, AA_perc:EA_perc, factor_key=TRUE) %>% 
  mutate(race=recode(race,`AA_perc`="AA",`EA_perc`="EA"))
```

```{r}
a<-ggplot(temp_ad4)+
  geom_point(aes(x=month,y=perc,col=race)) +
  scale_x_continuous(breaks = c(12,24,36,48,60))+
  scale_color_manual(values = c("AA" = cbPalette[2],
                                "EA" = cbPalette[3])) +
  labs(subtitle="A",x="", y="Censoring (%)",col="Race")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', 
                                        color = 'white'),
        legend.key=element_blank(),
        legend.position = "none",
        text=element_text(family="Times New Roman"))
a
```

### FA

```{r}
months <-0:59
fa_total<-fa_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(total=n())
t(fa_total)

temp_fa<-fa_total
for (i in months){
  cens<-fa_wide %>%
    group_by(PAT_MRN_ID) %>% 
    arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
    filter(row_number()==n()) %>% 
    filter(tstart>=i) %>% 
    ungroup() %>% 
    group_by(RACE) %>% 
    summarise("n.{i}":=n())
  
  temp_fa<-temp_fa %>%
    left_join(cens,by="RACE")
}
temp_fa
temp_fa2<-as.matrix(t(temp_fa))
temp_fa3<-data.frame(month=months,
                     AA=as.numeric(temp_fa2[3:nrow(temp_fa2),1]),
                     EA=as.numeric(temp_fa2[3:nrow(temp_fa2),2])
                     )
temp_fa3$AA_perc<-1-round(temp_fa3$AA/fa_total$total[1],2)
temp_fa3$EA_perc<-1-round(temp_fa3$EA/fa_total$total[2],2)
temp_fa4<-gather(temp_fa3, race, perc, AA_perc:EA_perc, factor_key=TRUE) %>% 
  mutate(race=recode(race,`AA_perc`="AA",`EA_perc`="EA"))
```

```{r}
b<-ggplot(temp_fa4)+
  geom_point(aes(x=month,y=perc,col=race)) +
  scale_x_continuous(breaks = c(12,24,36,48,60))+
  scale_color_manual(values = c("AA" = cbPalette[2],
                                "EA" = cbPalette[3])) +
  labs(subtitle="B",x="", y="",col="Race")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', 
                                        color = 'white'),
        legend.key=element_blank(),
        legend.position = "none",
        text=element_text(family="Times New Roman"))
b
```

### Asthma

```{r}
months <-0:59
as_total<-as_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(total=n())
t(as_total)

temp_as<-as_total
for (i in months){
  cens<-as_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  filter(tstart>=i) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise("n.{i}":=n())
  
  temp_as<-temp_as %>%
    left_join(cens,by="RACE")
}
temp_as
temp_as2<-as.matrix(t(temp_as))
temp_as3<-data.frame(month=months,
                     AA=as.numeric(temp_as2[3:nrow(temp_as2),1]),
                     EA=as.numeric(temp_as2[3:nrow(temp_as2),2])
                     )
temp_as3$AA_perc<-1-round(temp_as3$AA/as_total$total[1],2)
temp_as3$EA_perc<-1-round(temp_as3$EA/as_total$total[2],2)
temp_as4<-gather(temp_as3, race, perc, AA_perc:EA_perc, factor_key=TRUE) %>% 
  mutate(race=recode(race,`AA_perc`="AA",`EA_perc`="EA"))
```

```{r}
c<-ggplot(temp_as4)+
  geom_point(aes(x=month,y=perc,col=race)) +
  scale_x_continuous(breaks = c(12,24,36,48,60))+
  scale_color_manual(values = c("AA" = cbPalette[2],
                                "EA" = cbPalette[3])) +
  labs(subtitle="C",x="", y="Censoring (%)",col="Race")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', 
                                        color = 'white'),
        legend.key=element_blank(),
        legend.position = "none",
        text=element_text(family="Times New Roman"))
c
```


### AR

```{r}
months <-0:59
ar_total<-ar_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(total=n())
t(ar_total)

temp_ar<-ar_total
for (i in months){
  cens<-ar_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  filter(tstart>=i) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise("n.{i}":=n())
  
  temp_ar<-temp_ar %>%
    left_join(cens,by="RACE")
}
temp_ar
temp_ar2<-as.matrix(t(temp_ar))
temp_ar3<-data.frame(month=months,
                     AA=as.numeric(temp_ar2[3:nrow(temp_ar2),1]),
                     EA=as.numeric(temp_ar2[3:nrow(temp_ar2),2])
                     )
temp_ar3$AA_perc<-1-round(temp_ar3$AA/ar_total$total[1],2)
temp_ar3$EA_perc<-1-round(temp_ar3$EA/ar_total$total[2],2)
temp_ar4<-gather(temp_ar3, race, perc, AA_perc:EA_perc, factor_key=TRUE) %>% 
  mutate(race=recode(race,`AA_perc`="AA",`EA_perc`="EA"))
```

```{r}
d<-ggplot(temp_ar4)+
  geom_point(aes(x=month,y=perc,col=race)) +
  scale_x_continuous(breaks = c(12,24,36,48,60))+
  scale_color_manual(values = c("AA" = cbPalette[2],
                                "EA" = cbPalette[3])) +
  labs(subtitle="D",x="Time (months)", y="",col="Race")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', 
                                        color = 'white'),
        legend.key=element_blank(),
        legend.position = "none",
        text=element_text(family="Times New Roman"))
d
```

### EoE

```{r}
months <-0:59
eoe_total<-eoe_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(total=n())
t(eoe_total)

temp_eoe<-eoe_total
for (i in months){
  cens<-eoe_wide %>%
  group_by(PAT_MRN_ID) %>% 
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  filter(row_number()==n()) %>% 
  filter(tstart>=i) %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise("n.{i}":=n())
  
  temp_eoe<-temp_eoe %>%
    left_join(cens,by="RACE")
}
temp_eoe
temp_eoe2<-as.matrix(t(temp_eoe))
temp_eoe3<-data.frame(month=months,
                     AA=as.numeric(temp_eoe2[3:nrow(temp_eoe2),1]),
                     EA=as.numeric(temp_eoe2[3:nrow(temp_eoe2),2])
                     )
temp_eoe3$AA_perc<-1-round(temp_eoe3$AA/eoe_total$total[1],2)
temp_eoe3$EA_perc<-1-round(temp_eoe3$EA/eoe_total$total[2],2)
temp_eoe4<-gather(temp_eoe3, race, perc, AA_perc:EA_perc, factor_key=TRUE) %>% 
  mutate(race=recode(race,`AA_perc`="AA",`EA_perc`="EA"))
```


```{r}
e<-ggplot(temp_eoe4)+
  geom_point(aes(x=month,y=perc,col=race)) +
  scale_x_continuous(breaks = c(12,24,36,48,60))+
  scale_color_manual(values = c("AA" = cbPalette[2],
                                "EA" = cbPalette[3])) +
  labs(subtitle="E",x="Time (months)", y="Censoring (%)",col="Race")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', 
                                        color = 'white'),
        legend.key=element_blank(),
        legend.key.size = unit(6,"point"),
        legend.title = element_text(size=13),
        legend.text=element_text(size=13),
        text=element_text(family="Times New Roman"))
## Function to extract legend
get_legend <- function(a.gplot){ 
    tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
    legend <- tmp$grobs[[leg]] 
    legend
} 

legend<-get_legend(e)
e<-e+theme(legend.position = "none")


```


```{r}
f_censoring<-(a+b)/(c+d)/(e+legend)
f_censoring

ggsave(filename = "images/fs.censoring.png", f_censoring,# en cm mejor para mantener el tamanho
  width=2200,height = 2600, dpi = 300, units = "px", device='png')
```

