---
title: "Untitled"
author: "Esteban Correa"
date: "2022-08-04"
output: html_document
---

-v2 
updated to run Apple Silicon

-v3
code reorganized, 
Neat incidence curve using geom_textpath

v5
drop of deprivation index, EoE added to comorbidities

v6
Sankey diagram added again

v7
Use of atm_survival_v8.RData with improved inclusion criteria
t-test reemplazados por lmm para time-varying covariates
code clean-up
survival with formula f7
RMST figure removed 

v8
new survival with table

v9 
jaccard plot added to figure 3 panel B

v10
Include fixes on number at risk counts introduced by ggsurvminer
Use v12 models (new splines2 and rstanarm versions)

v11
x-axis rugs added for reviewer comments

v12
Use v13 models (centered covariates and frailty term)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(rstanarm)
library(survival)
library(tidyverse)
library(patchwork)
library(bayesplot)
library(loo)
library(ggalluvial)
library(ggraph)
library(geomtextpath)
library(tableone)

cbPalette <- c("#999999",  "#E64B35", "#4DBBD5", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

source("helpers.r")

load("data/atm_survival_v8.RData")
ad_wide$myICD<-"L20"
fa_wide$myICD<-"Z91"
as_wide$myICD<-"J45"
ar_wide$myICD<-"J30"
eoe_wide$myICD<-"K20"
all_diseases<-rbind(ad_wide,as_wide,ar_wide,fa_wide,eoe_wide)%>% 
  mutate(
    comorbidities=rowSums(across(c(AD,AS,AR,FA,EoE))),
    onemore=if_else(comorbidities>=2,"Yes","No")
  )%>% 
  filter(edvisit==1) %>% 
  ungroup() %>%
  mutate(PAT_MRN_ID2=as.numeric(PAT_MRN_ID),
    age_visit.cgm = c(scale(age_visit, scale = T)),
    ndvi.cgm = c(scale(ndvi, scale = T)),
    avgDailyPM.cgm = c(scale(avgDailyPM, scale = T))
  ) %>%
  group_by(RACE) %>%
  mutate(ndvi.cwc = c(scale(ndvi, scale = T)),
         avgDailyPM.cwc = c(scale(avgDailyPM, scale = T))) %>%
  ungroup()

```



```{r}


data.ad<-ad_wide %>%
  mutate(GENDER=factor(GENDER),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(FA,AS,AR,EoE))),
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>%
  arrange(PAT_MRN_ID,enum)

# Note: imputation of covariates for last visit is required to enable right censoring
data.as<-as_wide %>%
  mutate(GENDER=factor(GENDER),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AR,FA,EoE))), # except Asthma
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>% 
  arrange(PAT_MRN_ID,enum)

data.ar<-ar_wide %>%
  mutate(GENDER=factor(GENDER),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AS,FA,EoE))), #except rhinitis
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>% 
  arrange(PAT_MRN_ID,enum)

data.fa<-fa_wide %>%
  mutate(GENDER=factor(GENDER),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AS,AR,EoE))), #except FA
    onemore=if_else(comorbidities>=1,TRUE,FALSE) 
  )%>% 
  arrange(PAT_MRN_ID,enum)

data.eoe<-eoe_wide %>%
  mutate(GENDER=factor(GENDER),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AS,AR,FA))), #except eoe
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>% 
  arrange(PAT_MRN_ID,enum)


```






# Figure 1: 



## Incidence graph


```{r}
firstEvent.ad<-ad_wide %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(row_number()==1) %>% 
  mutate(ICD="AD")%>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  mutate(age_on=round(mean(age_visit),1))
  
firstEvent.as<-as_wide %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(row_number()==1) %>% 
  mutate(ICD="Asthma")%>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  mutate(age_on=round(mean(age_visit),1))

firstEvent.ar<-ar_wide %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(row_number()==1) %>% 
  mutate(ICD="AR")%>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  mutate(age_on=round(mean(age_visit),1))
         
firstEvent.fa<-fa_wide %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(row_number()==1) %>% 
  mutate(ICD="FA")%>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  mutate(age_on=round(mean(age_visit),1))

firstEvent.eoe<-eoe_wide %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(row_number()==1) %>% 
  mutate(ICD="EoE") %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  mutate(age_on=round(mean(age_visit),1))

firstEvent.all<-rbind(firstEvent.ad,firstEvent.as,firstEvent.ar,firstEvent.fa,firstEvent.eoe) %>% 
  mutate(ICD=factor(recode(ICD,`AS`="Asthma"),
                         levels = c("AD","FA", "Asthma", "AR","EoE"))
         ) 
```

plot version b

```{r}

race_names <- c(
  'AA'="African american",
  'EA'="European american"
)

# races horizontal
panelA<-ggplot(firstEvent.all, aes(x = age_visit, colour = ICD)) +
  geom_density(size = 0.8,key_glyph=draw_key_path) +
  # geom_textdensity(aes(hjust=ICD),size = 3.5, fontface = 2,vjust = 0.3) +
  scale_hjust_discrete()+
  labs(y="Incidence density of disease",x="Age of onset (years)",col="Allergic\nDisease",fill="",subtitle = "A") +
  theme_minimal()+
  # theme(legend.position = c(0.94,0.8),text=element_text(family="Times New Roman"))+
  theme(legend.position = "right",text=element_text(family="Times New Roman"))+
    facet_wrap(RACE~., 
             labeller=labeller(RACE=race_names)
             )
panelA



# ggsave(filename = "images/figure2a_v2.png", f2,# en cm mejor para mantener el tamanho
# width=2200,height = 1600, dpi = 300, units = "px", device='png')
```

version C separated

```{r}
# races horizontal
panelA.aa<-ggplot(firstEvent.all%>% 
  filter(RACE=="AA"), aes(x = age_visit, colour = ICD)) +
  geom_density(size = 0.8,key_glyph=draw_key_path) +
  # geom_textdensity(aes(hjust=ICD),size = 3.5, fontface = 2,vjust = 0.3) +
  scale_hjust_discrete()+ylim(0,0.3)+
  labs(y="Incidence density of disease",x="Age (years)",col="Allergic\nDisease",fill="",title="A",subtitle = "African American") +
  theme_minimal()+
  # theme(legend.position = c(0.94,0.8),text=element_text(family="Times New Roman"))+
  theme(legend.position = "right",text=element_text(family="Times New Roman"))
leg1<-get_legend(panelA.aa)
panelA.aa<-panelA.aa+theme(legend.position = "none")


panelA.ea<-ggplot(firstEvent.all%>% 
  filter(RACE=="EA"), aes(x = age_visit, colour = ICD)) +
  geom_density(size = 0.8,key_glyph=draw_key_path) +
  # geom_textdensity(aes(hjust=ICD),size = 3.5, fontface = 2,vjust = 0.3) +
  scale_hjust_discrete()+ylim(0,0.3)+
  labs(y="",x="Age (years)",col="Allergic\nDisease",fill="",title="",subtitle = "European American") +
  theme_minimal()+
  # theme(legend.position = c(0.94,0.8),text=element_text(family="Times New Roman"))+
  theme(legend.position = "right",text=element_text(family="Times New Roman"))
panelA<-panelA.aa+panelA.ea+ plot_layout(guides = "collect") 
panelA
```



# disease progression

Using dense rank to deal with ties
https://stackoverflow.com/questions/46981273/find-the-top-n-rows-with-ties-in-r-using-head


```{r}
#first diagnosis of each disease
diag10<-all_diseases %>%
  group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID,myICD,VISIT_DATE2) %>%
  distinct(myICD,.keep_all = TRUE)

#mark visits with duplicated dates
diag11<-diag10 %>%
  group_by(PAT_MRN_ID) %>%
  arrange(PAT_MRN_ID,VISIT_DATE2) %>% 
  mutate(isDup=duplicated(VISIT_DATE2) | duplicated(VISIT_DATE2, fromLast = TRUE),
         grp = dense_rank(VISIT_DATE2))

# We deal with ties at first and second visit
diag12<-diag11 %>%
  group_by(PAT_MRN_ID) %>%
  mutate(nVisits=n()) %>% 
  filter(nVisits >= 2) %>%
  # filter(grp <= 2) %>%
  filter(max(grp)==2) %>% 
  mutate(enum2=factor(paste0("visit","_",comorbidities))) 
```


we can build a graph 


```{r}
a <- c("A", "B")
b <- c("X", "Y", "Z")
# Use the outer function to apply the paste function to each pair of elements from a and b
c <- outer(a, b, paste, sep = "")
# Print the output vector c
print(c)

# Use the as.vector function to convert the matrix to a vector
v <- as.vector(c)
# Use the matrix function to convert the vector to a one column matrix
m1 <- matrix(v, ncol = 1)
# Print the reshaped matrix
print(m1)

a <- c("A", "B","C")
b <- c("X")
outer(a, b, paste, sep = "")
```

```{r}
# library(tidyr)
patoi<-diag12 %>% 
  distinct(PAT_MRN_ID) %>% 
  pull(PAT_MRN_ID)
fromtoDF<-{}

for (i in patoi){
  # print(i)
# curPat<-"0000694988"
# curPat<-"0000801595"
  tempDF<-filter(diag12,PAT_MRN_ID %in% i)
  aVec<-list()
  bVec<-list()
  for(j in 1:nrow(tempDF)){
    if (tempDF[j,]$grp==1){
      aVec<-append(aVec,tempDF[j,]$myICD)
    }else{
      bVec<-append(bVec,tempDF[j,]$myICD)
    }
  }
  # print(aVec)
  # print(bVec)
  c<-outer(aVec, bVec, paste, sep = "-")
  # print(c)
  d<-as.data.frame(matrix2column(c))
  # print(d)
  e<-d %>% separate(fromto, sep="-", c("from", "to"))
  e$PAT_MRN_ID<-i
  e$RACE<-first(unique(tempDF$RACE))
  # print(e)
  fromtoDF<-rbind(fromtoDF,e)
}
```

```{r}
fromtoDF2<-fromtoDF %>%
  mutate(from=as.factor(recode(from,
                         `J30`="AR",
                         `J45`="Asthma",
                         `K20`="EoE",
                         `L20`="AD",
                         `Z91`="FA")),
         to=as.factor(recode(to,
                         `J30`="AR",
                         `J45`="Asthma",
                         `K20`="EoE",
                         `L20`="AD",
                         `Z91`="FA")),
         )

```


Using summary

```{r}

fromtoDF2Sum<-fromtoDF2 %>% 
  group_by(from,to,RACE) %>% 
  summarize(freq=n()) %>% 
  mutate(arc_direction=ifelse(RACE=="AA",-1,1),
         from=factor(from, ordered = TRUE, 
                                levels = c("AD","FA","Asthma","AR","EoE")),
         to=factor(to, ordered = TRUE, 
                                levels = c("AD","FA","Asthma","AR","EoE")),
         )


```





## multimorbidity proportion by race

```{r}
multim.df.aa<-all_diseases %>% 
  ungroup() %>% 
  filter(RACE=="AA") %>% 
  select(c("AD","FA","AS","AR","EoE"))

multim.df.ea<-all_diseases %>% 
  ungroup() %>% 
  filter(RACE=="EA") %>% 
  select(c("AD","FA","AS","AR","EoE"))

multi.mat.aa<-get_prop_matrix(as.matrix(multim.df.aa),NULL)
colnames(multi.mat.aa$mat.jaccard)<-c("AD","FA","Asthma","AR","EoE")
rownames(multi.mat.aa$mat.jaccard)<-c("AD","FA","Asthma","AR","EoE")
multi.mat.aa

multi.mat.ea<-get_prop_matrix(as.matrix(multim.df.ea),1)
colnames(multi.mat.ea$mat.jaccard)<-c("AD","FA","Asthma","AR","EoE")
rownames(multi.mat.ea$mat.jaccard)<-c("AD","FA","Asthma","AR","EoE")
multi.mat.ea
```

 


```{r}
upper_tri.aa <- get_upper_tri(multi.mat.aa$mat.jaccard)
upper_tri.ea <- get_upper_tri(multi.mat.ea$mat.jaccard)

#long format
melted_aa <- melt(upper_tri.aa, na.rm = TRUE)
melted_ea <- melt(upper_tri.ea, na.rm = TRUE)

aa<-ggplot(data = melted_aa, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
  # scale_fill_distiller(palette = "YlOrRd",direction = 1)+
  scale_fill_gradient2(low = "lightblue", high = "red", mid = "orange", 
   midpoint = 0.5, limit = c(0,1), space = "Lab", 
   name="Co-Occurrence\nProportion (%)")+
  theme_minimal()+ 
  labs(title="B",subtitle="African American")+
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()+
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4,family="Times New Roman") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    text=element_text(family="Times New Roman")
    # legend.position = "bottom",
    # legend.direction = "horizontal"
    )+
    guides(fill = guide_colorbar(barwidth = 1, barheight = 7,
                  title.position = "top", title.hjust = 0.5))
aa
leg<-get_legend(aa)
aa<-aa+theme(legend.position = "none")

ea<-ggplot(data = melted_ea, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
  # scale_fill_distiller(palette = "YlOrRd",direction = 1)+
  scale_fill_gradient2(low = "lightblue", high = "red", mid = "orange", 
   midpoint = 0.5, limit = c(0,1), space = "Lab", 
   name="Jaccard\nIndex")+
  theme_minimal()+ 
  labs(title="", subtitle="European American")+
  coord_fixed()+
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4,family="Times New Roman") +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    text=element_text(family="Times New Roman"),
    legend.position = "right",
    # legend.direction = "horizontal"
    )+
    guides(fill = guide_colorbar(barwidth = 1, barheight = 7,
                  title.position = "top", title.hjust = 0.5))
ea

panelB<-aa+ea+plot_layout(guides = "collect") 
panelB
```


```{r}
# f2<-panelA/panelB
 # plot_layout(guides = "collect") 
f2<-(panelA.aa/aa)|(panelA.ea/ea)

f2
ggsave(filename = "images/f2_v9.png", f2,# en cm mejor para mantener el tamanho
  width=2200,height = 2600, dpi = 300, units = "px", device='png')
ggsave(filename = "images/f2_v9.tif", f2,# en cm mejor para mantener el tamanho
  width=2200,height = 2600, dpi = 300, units = "px", device='tiff',compression="lzw")
```


##Table S1

```{r}
fromtoAD<-fromtoDF2 %>% 
  filter(from == "AD") %>% 
  mutate(to=factor(to, ordered = TRUE, 
                                levels = c("AD","FA","Asthma","AR","EoE"))) %>% 
  mutate(to=droplevels(to))

myVars<-c("to")
tab <- CreateTableOne(vars = myVars,
                      # strata = "RACE" ,
                      data = fromtoAD,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = fromtoAD,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
```


```{r}
fromtoFA<-fromtoDF2 %>% 
  filter(from == "FA") %>% 
  mutate(to=factor(to, ordered = TRUE, 
                   levels = c("AD","FA","Asthma","AR","EoE"))) %>% 
  mutate(to=droplevels(to))

tab <- CreateTableOne(vars = myVars,
                      # strata = "RACE" ,
                      data = fromtoFA,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = fromtoFA,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 
```


```{r}
fromtoAS<-fromtoDF2 %>% 
  filter(from == "Asthma") %>% 
  mutate(to=factor(to, ordered = TRUE, 
                   levels = c("AD","FA","Asthma","AR","EoE"))) %>% 
  mutate(to=droplevels(to))

tab <- CreateTableOne(vars = myVars,
                      # strata = "RACE" ,
                      data = fromtoAS,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 

tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = fromtoAS,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 
```

```{r}
library(ggraph)
library(igraph)
fromtoAR<-fromtoDF2 %>% 
  filter(from == "AR") %>% 
  mutate(to=factor(to, ordered = TRUE, 
                   levels = c("AD","FA","Asthma","AR","EoE"))) %>% 
  mutate(to=droplevels(to))

tab <- CreateTableOne(vars = myVars,
                      # strata = "RACE" ,
                      data = fromtoAR,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 

tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = fromtoAR,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 
```


```{r}
fromtoEoE<-fromtoDF2 %>% 
  filter(from == "EoE") %>% 
  mutate(to=factor(to, ordered = TRUE, 
                   levels = c("AD","FA","Asthma","AR","EoE"))) %>% 
  mutate(to=droplevels(to))

tab <- CreateTableOne(vars = myVars,
                      # strata = "RACE" ,
                      data = fromtoEoE,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 

tab <- CreateTableOne(vars = myVars,
                      strata = "RACE" ,
                      data = fromtoEoE,
                       includeNA = T,
                          test = T)
tab2Mat<-print(tab, showAllLevels = TRUE, formatOptions = list(big.mark = ","),quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
tab2Mat%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F,font="Times New Roman") 
```





## Posterior uncertainty intervals

```{r}
load("models/ad.model.ms10.f9.v13.RData")
load("models/fa.model.ms10.f9.v13.RData")
load("models/as.model.ms10.f9.v13.RData")
load("models/ar.model.ms10.f9.v13.RData")
load("models/eoe.model.ms10.f9.v13.RData")
```

## Correlation

```{r}
dataCor<-all_diseases %>% 
  ungroup() %>% 
  dplyr::select(avgDailyPM,ndvi) %>% 
  drop_na()

r<-round(cor(dataCor[,"avgDailyPM"],dataCor[,"ndvi"],use = "complete.obs"),3)
cor.test(dataCor$avgDailyPM, dataCor$ndvi)
p<-cor.test(dataCor$avgDailyPM, dataCor$ndvi)$p.value


ggplot(all_diseases,aes(x=avgDailyPM,y=ndvi))+
  geom_point(size=0.1,alpha=0.1)+
  geom_smooth(method="lm", col="blue",se = T) +
  labs(x="PM2.5 (μm)",y="Greenspace (ndvi)")+
  annotate("text", x=18, y=0.6, label=paste0("r = ", r), hjust=0) +
  annotate("text", x=18, y=0.58, label=paste0("p = ", round(p, 3)), hjust=0) +
  theme_classic() 
```


##Last event

```{r}
fit.stan.ad.ms10.f7$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  slice_tail() %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(n=n())%>% 
  mutate(total=sum(n))%>% 
  mutate(prop = 100 * n / sum(n)) %>% 
  mutate(prop2=round(prop,1),
         prop2text=paste0(n," (",prop2,")"))

fit.stan.fa.ms10.f7$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  slice_tail() %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(n=n())%>% 
  mutate(total=sum(n))%>% 
  mutate(prop = 100 * n / sum(n)) %>% 
  mutate(prop2=round(prop,1),
         prop2text=paste0(n," (",prop2,")"))

fit.stan.as.ms10.f7$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  slice_tail() %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(n=n())%>% 
  mutate(total=sum(n)) %>% 
  mutate(prop = 100 * n / sum(n)) %>% 
  mutate(prop2=round(prop,1),
         prop2text=paste0(n," (",prop2,")"))

fit.stan.ar.ms10.f7$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  slice_tail() %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(n=n())%>% 
  mutate(total=sum(n)) %>% 
  mutate(prop = 100 * n / sum(n)) %>% 
  mutate(prop2=round(prop,1),
         prop2text=paste0(n," (",prop2,")"))

fit.stan.eoe.ms10.f7$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  slice_tail() %>% 
  ungroup() %>% 
  group_by(RACE) %>% 
  summarise(n=n()) %>% 
  mutate(prop = 100 * n / sum(n)) %>% 
  mutate(prop2=round(prop,1),
         prop2text=paste0(n," (",prop2,")"))

```


## Hazard ratio

### Follow up examination

```{r}
fu.ad_df<-fit.stan.ad.ms10.f9$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(edvisit==1) %>% 
  slice_tail() %>% 
  ungroup()
summary(fu.ad_df$tstart2)
hist(fu.ad_df$tstart2,32)

fu.fa_df<-fit.stan.fa.ms10.f9$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(edvisit==1) %>% 
  slice_tail() %>% 
  ungroup() 
summary(fu.fa_df$tstart2)

fu.as_df<-fit.stan.as.ms10.f9$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(edvisit==1) %>% 
  slice_tail() %>% 
  ungroup() 
summary(fu.as_df$tstart2)

fu.ar_df<-fit.stan.ar.ms10.f9$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(edvisit==1) %>% 
  slice_tail() %>% 
  ungroup() 
summary(fu.ar_df$tstart2)

fu.eoe_df<-fit.stan.eoe.ms10.f9$data %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(enum) %>% 
  filter(edvisit==1) %>% 
  slice_tail() %>% 
  ungroup() 
summary(fu.eoe_df$tstart2)
```

## Hazard baseline

```{r}
fits_stan <- list(
                  "AD" = fit.stan.ad.ms10.f9,
                  "FA" = fit.stan.fa.ms10.f9,
                  "Asthma" = fit.stan.as.ms10.f9,
                  "AR" = fit.stan.ar.ms10.f9,
                  "EoE" = fit.stan.eoe.ms10.f9
                  )
plots <- map(fits_stan,plot)

# v11 now includes geom_rug and follow-up stats
rug_df <-
  fit.stan.ad.ms10.f9$data %>% 
  mutate(edvisit = ifelse(edvisit == 1, tstart2, NA))
subt.lbl<-paste0("A\n Min: ",round(min(fu.ad_df$tstart2),1),
       ", Mean: ",round(mean(fu.ad_df$tstart2),1),
       ", Max: ",round(max(fu.ad_df$tstart2),1))
a<-plots[[1]]+
  labs(subtitle = subt.lbl,x="")+
  coord_cartesian(ylim = c(0,.7))+
  geom_rug(data=rug_df,aes(x = edvisit), sides = "b", alpha = 0.5,linewidth=0.1)+
  theme(plot.title = element_text(hjust = .0),
        plot.subtitle = element_text(margin=margin(t=10,b=-12)),
        text=element_text(family="Times New Roman"))

rug_df <-
  fit.stan.fa.ms10.f9$data %>% 
  mutate(edvisit = ifelse(edvisit == 1, tstart2, NA))
subt.lbl<-paste0("B\n Min: ",round(min(fu.fa_df$tstart2),1),
       ", Mean: ",round(mean(fu.fa_df$tstart2),1),
       ", Max: ",round(max(fu.fa_df$tstart2),1))
b<-plots[[2]]+
  labs(subtitle = subt.lbl,x="",y="")+
  coord_cartesian(ylim = c(0,1))+
  geom_rug(data=rug_df,aes(x = edvisit), sides = "b", alpha = 0.5,linewidth=0.1)+
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(margin=margin(t=10,b=-12)),
        text=element_text(family="Times New Roman"))

rug_df <-
  fit.stan.as.ms10.f9$data %>% 
  mutate(edvisit = ifelse(edvisit == 1, tstart2, NA))
subt.lbl<-paste0("C\n Min: ",round(min(fu.as_df$tstart2),1),
       ", Mean: ",round(mean(fu.as_df$tstart2),1),
       ", Max: ",round(max(fu.as_df$tstart2),1))
c<-plots[[3]]+
  labs(subtitle = subt.lbl,x="")+
  coord_cartesian(ylim = c(0,.9))+
  geom_rug(data=rug_df,aes(x = edvisit), sides = "b", alpha = 0.5,linewidth=0.1)+
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(margin=margin(t=10,b=-12)),
        text=element_text(family="Times New Roman"))

rug_df <-
  fit.stan.ar.ms10.f9$data %>% 
  mutate(edvisit = ifelse(edvisit == 1, tstart2, NA))
subt.lbl<-paste0("D\n Min: ",round(min(fu.ar_df$tstart2),1),
       ", Mean: ",round(mean(fu.ar_df$tstart2),1),
       ", Max: ",round(max(fu.ar_df$tstart2),1))
d<-plots[[4]]+labs(subtitle = subt.lbl,y="", x="Time (months)")+
  coord_cartesian(ylim = c(0,1))+
  geom_rug(data=rug_df,aes(x = edvisit), sides = "b", alpha = 0.5,linewidth=0.1)+
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(margin=margin(t=10,b=-12)),
        text=element_text(family="Times New Roman"))

rug_df <-
  fit.stan.eoe.ms10.f9$data %>% 
  mutate(edvisit = ifelse(edvisit == 1, tstart2, NA))
subt.lbl<-paste0("E\n Min: ",round(min(fu.eoe_df$tstart2),1),
       ", Mean: ",round(mean(fu.eoe_df$tstart2),1),
       ", Max: ",round(max(fu.eoe_df$tstart2),1))
e<-plots[[5]]+
  labs(subtitle = subt.lbl, x="Time (months)")+
  coord_cartesian(ylim = c(0,.8))+
  geom_rug(data=rug_df,aes(x = edvisit), sides = "b", alpha = 0.5,linewidth=0.1)+
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(margin=margin(t=10,b=-12)),
        text=element_text(family="Times New Roman"))

design <- "
  12
  34
  5#
"
fs1<-a+b+c+d+e+plot_layout(design = design)
fs1
ggsave(filename = "images/fs1_basehr_v3.png", fs1,# en cm mejor para mantener el tamanho
  width=2200,height = 2600, dpi = 300, units = "px", device='png')

```






# Survival

## Figure 1 Survival curves:AS vs AR vs FA

```{r}
RiskSetCount <- function(timeindex, patientsdf,strataoi) {
  atrisk <- NULL
  survivaltime<-patients$tstop2[patients$Strata==strataoi]
  for (t in timeindex)
    atrisk <- c(atrisk, sum(survivaltime >= t))
  df<-data.frame(strata=strataoi,value=atrisk,time=timeindex)
  return(df)
}
```

Data generation process

```{r}
data_test<-data.frame(id=1:4,
                      RACE = factor(rep(c("AA","EA"),each=2),levels=c("EA","AA")),
                      hcType=factor("Medicaid",levels=c("Commercial","Medicaid")),
                      age_visit=median(all_diseases$age_visit,na.rm = T),
                      avgDailyPM=median(all_diseases$avgDailyPM,na.rm = T),
                      ndvi=median(all_diseases$ndvi,na.rm = T),
                      age_visit.cgm=median(all_diseases$age_visit.cgm,na.rm = T),
                      avgDailyPM.cgm=median(all_diseases$avgDailyPM.cgm,na.rm = T),
                      ndvi.cgm=median(all_diseases$ndvi.cgm,na.rm = T),
                      onemore=rep(c(F,T),2),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      ) %>%
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
data_test
ndraws=1000
```


```{r}
annoTextS=4
# cbPalette <- c("#E64B35","#F0E442", "#4DBBD5", "#009E73")
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)


# already collapsed
# summary(fit.stan.ad.ms10.f9)
ps<-posterior_survfit(fit.stan.ad.ms10.f9,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA=1.18 (1.06 to 1.33)",
          "HR EA Comorbid EA=1.12 (1.02 to 1.23)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "none")
################ Risk table as ggplot element
datatr <- fit.stan.ad.ms10.f9$data %>%   ungroup() %>%
  # select(RACE,onemore) %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```

```{r}
mcmc_post_ci(fit.stan.fa.ms10.f9,.95,nvar = 12)%>%
  kableExtra::kbl() %>%
  kableExtra::kable_paper("hover",full_width=F)
```



```{r}
# FA
# already collapsed
ps<-posterior_survfit(fit.stan.fa.ms10.f9,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  RACE="AA",
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA Comorbid=1.35 (1.21 to 1.50)",
          "HR EA Comorbid=1.48 (1.40 to 1.58)")
)
################ survival curves
b<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="B",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
b<-b+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "none",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
b
################ Risk table as ggplot element
datatr <- fit.stan.fa.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  # scale_color_manual(values=cbPalette)+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))
(b<-b / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```



```{r}
######AS
# already collapsed
ps<-posterior_survfit(fit.stan.as.ms10.f9,
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
#tidybayes does  not work with posterior_survfit yet
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85,0.75),
  label=c("HR AA=1.16 (1.07 to 1.25)",
          "HR AA Comorbid=1.42 (1.35 to 1.51)",
          "HR EA Comorbid=1.56 (1.48 to 1.64)")
)
################ survival curves
c<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="C",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
c<-c+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "none",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
c
################ Risk table as ggplot element
datatr <- fit.stan.as.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))
(c<-c / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```


```{r}
mcmc_post_ci(fit.stan.ar.ms10.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
```


```{r}
#####AR
# already collapsed
ps<-posterior_survfit(fit.stan.ar.ms10.f9,
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
#tidybayes does  not work with posterior_survfit yet
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA Comorbid=1.62 (1.52 to 1.72)",
          "HR EA Comorbid=1.55 (1.46 to 1.64)")
)
################ survival curves
d<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="D",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
d<-d+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "none",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
d
################ Risk table as ggplot element
datatr <- fit.stan.ar.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))
(d<-d / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```



```{r}
#EoE
# already collapsed
ps<-posterior_survfit(fit.stan.eoe.ms10.f9,
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
#tidybayes does  not work with posterior_survfit yet
text.df<-data.frame(
  x=c(15),
  y=c(0.95),
  label=c(
          "HR EA Comorbid=1.21 (1.03 to 1.41)")
)
################ survival curves
e<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="E",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
e<-e+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "none",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
e
################ Risk table as ggplot element
datatr <- fit.stan.eoe.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))
(e<-e / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```



```{r}
f4.final<-ggpubr::ggarrange(a,b,c,d,e,legend,heights = c(1,1,1,1,1,2),
          ncol = 2, nrow = 3)

ggsave(filename = "images/figure4.surv.v13.png", f4.final,
       # en cm mejor para mantener el tamanho
       width=3200,height = 4800, dpi = 300, units = "px", device='png')
```


# TTR

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#F0E442", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test <- data.frame(
  id = 1:4,
  RACE = rep(c("AA", "EA"), each = 2),
  hcType = "Medicaid",
  age_visit = rep(median(all_diseases$age_visit, na.rm = T), 2),
  avgDailyPM = median(all_diseases$avgDailyPM, na.rm = T),
  ndvi = median(all_diseases$ndvi, na.rm = T),
  onemore = rep(c(F, T), 2),
  GENDER = rep("Male", 2)
) %>%
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
# summary(fit.stan.ad.ms10.f9)
ps<-posterior_survfit(fit.stan.ad.ms10.f7,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR Comorbid AA=1.15 (1.09 to 1.21)",
          "HR Comorbid EA=1.46 (1.37 to 1.55)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata), 
              # fill = "gray90",
              alpha=0.2,
              size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "none")
################ Risk table as ggplot element
datatr <- fit.stan.ad.ms10.f7$data %>%   ungroup() %>%
  # select(RACE,onemore) %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```
