---
title: "Untitled"
author: "Esteban Correa"
date: "2022-08-04"
output: html_document
---

-v2 
bayes for asthma dataset

-v3
Make use of v5 dataset
Generalization of rmst helper functions

-v4
Use onemore for one or more comordities AD/AS/AR/FA instead of Atopic dermatitis [Jocelyn's PARS paper].

-v5
Updated to run on Apple Silicon 

v6
data/atm_survival_v7.RData with public insurance added

v7 
drop of deprivation index, EoE added to comorbidities

v8
elpd, cindex, and calibration plots added

v9
Use of atm_survival_v8.RData with improved inclusion criteria
code clean up, 10k MCMC, 95% CI for HR and RMST
use of formula f6

v10
use of formula f7

v11
Use of atm_survival_v8.RData 
use of f7 formula
Add f0 for null model
Code clean up and f4 models removed

v12
Re run for new splines2 v0.5.1 and rstanarm v2.26.1

v13
make use centered versions for greenspace and pm2.5
Inclusion of frailty term in f9 equations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(rstanarm)
library(survival)
library(tidyverse)
library(patchwork)
library(bayesplot)
library(loo)
library(gtsummary)

source("helpers.r")
load("output/cchmc.as.v13.RData")
```



# Asthma readmission

## Dataset preparation


```{r}
load("data/atm_survival_v8.RData")

# Note: imputation of covariates for last visit is required to enable right censoring
data<-as_wide %>%
  mutate(GENDER=factor(GENDER,levels = c("Female", "Male")),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AR,FA,EoE))), # except Asthma
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>% 
  select(PAT_MRN_ID,tstart,tstop,tstart2,tstop2,edvisit,futime,enum,age_visit,GENDER,RACE,AD,FA,AS,AR,EoE,onemore,hcType,dep_index,ndvi,avgDailyPM) %>% 
  drop_na() %>% # added to avoid NA at validation
  arrange(PAT_MRN_ID,enum)

summary(as_wide)
summary(data)

data2<-data %>% filter(tstart2==tstop2)
data<-data %>% 
  filter(!(PAT_MRN_ID %in% data2$PAT_MRN_ID))

data %>% 
  distinct(PAT_MRN_ID,.keep_all = T) %>% 
  group_by(RACE) %>% 
  summarise(freq=n())

#Centering
data <- data %>%
  ungroup() %>%
  mutate(PAT_MRN_ID2=as.numeric(PAT_MRN_ID),
    age_visit.cgm = c(scale(age_visit, scale = T)),
    ndvi.cgm = c(scale(ndvi, scale = T)),
    avgDailyPM.cgm = c(scale(avgDailyPM, scale = T))
  ) %>%
  group_by(RACE) %>%
  mutate(ndvi.cwc = c(scale(ndvi, scale = T)),
         avgDailyPM.cwc = c(scale(avgDailyPM, scale = T))) %>%
  ungroup()
```


```{r}
#####use of tstart2,tstop2
f0=formula(Surv(tstart2, tstop2, edvisit) ~ 1)
cox.as0<-coxph(formula = f0,data =data)

# f5=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit*GENDER+RACE+hcType+onemore+ndvi+avgDailyPM,type="interval")
# cox.as5<-coxph(formula = f5,data =data)
# # summary(cox.as5)
# 
# #use of tstart2,tstop2
# f6=formula(Surv(tstart2, tstop2, edvisit) ~GENDER:age_visit+RACE+hcType+onemore+ndvi+avgDailyPM,type="interval")
# cox.as6<-coxph(formula = f6,data =data)
# # summary(cox.as6)

# use of interaction only for ndvi
f7=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi+RACE:avgDailyPM,type="interval")
cox.as7<-coxph(formula = f7,data =data)
summary(cox.as7)

# use of centered covariates
f8=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit.cgm:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm,type="interval")
cox.as8<-coxph(formula = f8,data =data)
summary(cox.as8)

# use of centered covariates  and frailty
f9=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit.cgm:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+frailty(PAT_MRN_ID),type="interval")
cox.as9<-coxph(formula = f9,data =data)
summary(cox.as9)

# use age*comorbidities interaction
f10=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit.cgm:GENDER+onemore:age_visit.cgm+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+frailty(PAT_MRN_ID),type="interval")
cox.as10<-coxph(formula = f10,data =data)
summary(cox.as10)


AIC(cox.as0,cox.as7,cox.as8,cox.as9,cox.as10)

cox.as7%>%
  tbl_regression(exponentiate = TRUE)
cox.as8%>%
  tbl_regression(exponentiate = TRUE)
cox.as9%>%
  tbl_regression(exponentiate = TRUE)
cox.as10%>%
  tbl_regression(exponentiate = TRUE)
```

## reversing beta coefficients

```{r}
logLik(cox.as7)
logLik(cox.as8) 
rescale.coefs <- function(beta,mu,sigma) {
    beta2 <- beta ## inherit names etc.
    # beta2[-1] <- sigma[1]*beta[-1]/sigma[-1]
    # cox
    beta2<- beta/sigma
    # beta2[1]  <- sigma[1]*beta[1]+mu[1]-sum(beta2[-1]*mu[-1])
    beta2
} 
```

Getting back centered coefficients in cox

```{r}
cox.as7$coefficients
cox.as8$coefficients

unsc.coeff <- cox.as7$coefficients[c(
  "age_visit:GENDERFemale",
  "age_visit:GENDERMale",
  "RACEEA:ndvi",
  "RACEAA:ndvi",
  "RACEEA:avgDailyPM",
  "RACEAA:avgDailyPM"
)]

scal.coeff <- cox.as8$coefficients[c(
  "age_visit.cgm:GENDERFemale",
  "age_visit.cgm:GENDERMale",
  "RACEEA:ndvi.cgm",
  "RACEAA:ndvi.cgm",
  "RACEEA:avgDailyPM.cgm",
  "RACEAA:avgDailyPM.cgm"
)]

unsc.var<-subset(data,select=c(age_visit,age_visit,ndvi,ndvi,avgDailyPM,avgDailyPM))
(cm <- colMeans(unsc.var))
(csd <- apply(unsc.var,2,sd))

exp(unsc.coeff)
exp(cc <-rescale.coefs(scal.coeff,mu=c(cm),sigma=c(csd)))
```



## Stan version 

```{r}

(f9mcmc=formula(Surv(tstart2, tstop2, edvisit) ~GENDER:age_visit.cgm+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+(1|PAT_MRN_ID2)))
# when chains>1 r makes use of viewer
CHAINS <- 4
CORES <- 4
ITER <- 1500
SEED <- 42
# draw from the prior predictive distribution of the stan_surv survival model
prior.stan.as <- stan_surv(
  formula = f9mcmc,
  data = data,
  basehaz = "exp",
  prior_PD = TRUE,
  iter = ITER,refresh=1500,
  chains = CHAINS,
  cores = CORES,
  seed = SEED)

prior.stan.as2 <- update(prior.stan.as,
                            prior_intercept = normal(0, 1),
                            prior = normal(0, .5))

prior.stan.as3 <- update(prior.stan.as,
                            prior_intercept = cauchy(0,10),
                            prior = normal(0,.3))

prior.stan.as4 <- update(prior.stan.as,
                            prior_intercept = cauchy(0,25),
                            prior = normal(0,.3))
prior.stan.as5 <- update(prior.stan.as,
                            prior_intercept = cauchy(0,10),
                            prior = normal(0,.5))
prior.stan.as6 <- update(prior.stan.as,
                            prior_intercept = cauchy(0,20),
                            prior = normal(0,.5))

```

improving priors:

```{r}
# extract HR from classical coxph for arm=B
exp(coef(cox.as9))[1]
base_cox_hr <- vline_at(exp(coef(cox.as9))[1], color = "green")
a<-mcmc_hist(prior.stan.as,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors default")
b<-mcmc_hist(prior.stan.as2,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors N(0,1),N(0,.5))")
c<-mcmc_hist(prior.stan.as3,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors C(10),N(0,.3))")
d<-mcmc_hist(prior.stan.as4,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors C(25),N(0,.3))")
e<-mcmc_hist(prior.stan.as4,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors C(10),N(0,.5))")
f<-mcmc_hist(prior.stan.as4,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors C(25),N(0,.5))")

a+b+c+d+e+f+plot_layout(ncol = 3)
```


## Sampling exp+mspline

```{r}
f0
f8mcmc=formula(Surv(tstart2, tstop2, edvisit) ~GENDER:age_visit.cgm+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm)
(f9mcmc=formula(Surv(tstart2, tstop2, edvisit) ~GENDER:age_visit.cgm+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+(1|PAT_MRN_ID2)))
# exp
fit.stan.as.exp.f7 <- update(prior.stan.as2,  
                             prior_PD = FALSE,
                             formula=f7,
                             basehaz = "exp")

fit.stan.as.exp.f8 <- update(prior.stan.as2,  
                             prior_PD = FALSE,
                             formula=f8mcmc,
                             basehaz = "exp")

fit.stan.as.exp.f9 <- update(prior.stan.as2,  
                             prior_PD = FALSE,
                             formula=f9mcmc,
                             basehaz = "exp")

# cubic m-spline tstart2, tstop2
fit.stan.as.ms10.f7 <- update(fit.stan.as.exp.f7,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))

fit.stan.as.ms10.f8 <- update(fit.stan.as.exp.f8,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))
fit.stan.as.ms10.f9 <- update(fit.stan.as.exp.f9,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))

fit.stan.as.exp.f0 <- update(prior.stan.as2,  
                             prior_PD = FALSE,
                             data=fit.stan.as.ms10.f8$data,
                             formula=f0,
                             basehaz = "exp")

fits_stan <- list("exp.f0" = fit.stan.as.exp.f0,
                  "exp.f7" = fit.stan.as.exp.f7,
                  "ms10.f7" = fit.stan.as.ms10.f7,
                  "exp.f8" = fit.stan.as.exp.f8,
                  "ms10.f8" = fit.stan.as.ms10.f8,
                  "exp.f9" = fit.stan.as.exp.f9,
                  "ms10.f9" = fit.stan.as.ms10.f9
                  )

#Are chains mixing well?
color_scheme_set("mix-blue-red")
mcmc_trace(fit.stan.as.exp.f9, 
           pars=c("RACEAA", "(Intercept)"),
           facet_args = list(ncol = 1, strip.position = "left")
           )
```


## Posterior uncertainty intervals

```{r}
mcmc_post_ci(fit.stan.as.exp.f7,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

mcmc_post_ci(fit.stan.as.ms10.f7,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

mcmc_post_ci(fit.stan.as.exp.f8,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

mcmc_post_ci(fit.stan.as.ms10.f8,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

mcmc_post_ci(fit.stan.as.exp.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

mcmc_post_ci(fit.stan.as.ms10.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
```


## Hazard ratio checking

```{r}
# extract HR from classical coxph for arm=B
exp(coef(cox.as9))[1]
base_cox_hr <- vline_at(exp(coef(cox.as9))[1], color = "green")

a<-mcmc_hist(prior.stan.as2,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle="Priors N(0,1),N(0,.5)")
b<-mcmc_hist(fit.stan.as.exp.f7,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f7 (exp)")
c<-mcmc_hist(fit.stan.as.ms10.f7,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f7 (ms-10)")
d<-mcmc_hist(fit.stan.as.exp.f8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f8 (exp)")
e<-mcmc_hist(fit.stan.as.ms10.f8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f8 (ms-10)")
f<-mcmc_hist(fit.stan.as.exp.f9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f9 (exp)")
g<-mcmc_hist(fit.stan.as.ms10.f9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f9 (ms-10)")

a+b+c+d+e+f+g+plot_layout(ncol = 4)
```



## Hazard baseline

```{r}
plots <- map(fits_stan,plot)

a<-plots[[1]]+
  labs(title = "Null")+
  coord_cartesian(ylim = c(0,.5))+
  theme(plot.title = element_text(hjust = .5))

b<-plots[[2]]+labs(title = "f8 (exp)")+
  coord_cartesian(ylim = c(0,.5))+
  theme(plot.title = element_text(hjust = .5))

c<-plots[[3]]+labs(title = "f8  (ms10)")+
  coord_cartesian(ylim = c(0,.9))+
  theme(plot.title = element_text(hjust = .5))

d<-plots[[4]]+labs(title = "f9 (exp)")+
  coord_cartesian(ylim = c(0,.5))+
  theme(plot.title = element_text(hjust = .5))

e<-plots[[5]]+labs(title = "f9  (ms10)")+
  coord_cartesian(ylim = c(0,.9))+
  theme(plot.title = element_text(hjust = .5))
a+b+c+d+e
```


# Survival curves: Posterior vs COX PH

Preparing synthetic dataset:

```{r}
# f9 centered+frailty
data_test<-data.frame(id=1:4,
                      RACE = factor(rep(c("AA","EA"),each=2),levels=c("EA","AA")),
                      hcType=factor("Medicaid",levels=c("Commercial","Medicaid")),
                      age_visit=rep(median(data$age_visit,na.rm = F),2),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=rep(median(data$age_visit.cgm,na.rm = F),2),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=rep(c(F,T),2),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )
```

##f7 plot

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test<-data_test %>% 
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
ps<-posterior_survfit(fit.stan.as.ms10.f7,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA Comorbid=2.09 (2.02 to 2.16)",
          "HR EA Comorbid=2.28 (2.21 to 2.35)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "right")
################ Risk table as ggplot element
datatr <- fit.stan.as.ms10.f7$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "right",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```


##f8 plot

```{r}
# annoTextS=4
# cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
# grid <- seq(0,60,by=12)
# 
# data_test<-data_test %>% 
#   mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
# ndraws=500
# # already collapsed
# ps<-posterior_survfit(fit.stan.as.ms10.f8,  
#                       newdata = data_test,
#                       times = 0,
#                       extrapolate   = T, 
#                       condition     = FALSE,
#                       return_matrix = F,
#                       control = list(edist = 60),
#                       draws = ndraws)
# ps<-ps %>% 
#   left_join(data_test,by="id")
# # prepare HR annotations
# text.df<-data.frame(
#   x=c(15),
#   y=c(0.95,0.85),
#   label=c("HR AA Comorbid=1.38 (1.09 to 1.74)",
#           "HR EA Comorbid=1.63 (1.49 to 1.78)")
# )
# ################ survival curves
# a<-ps %>%as_tibble()%>%
#   ggplot(aes(x=time,y=median,col=Strata)) +
#   # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
#   #             # fill = "gray90",
#   #             alpha=0.2,
#   #             size=0.0) +
#   geom_line()+
#   scale_color_manual(values=cbPalette)+
#   scale_fill_manual(values = cbPalette)+
#   scale_x_continuous(breaks = grid)+
#   labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
#   survminer::theme_survminer(base_family = "Times New Roman")
# a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
#            size=annoTextS,
#            hjust=0,family="Times New Roman")+
#   theme(legend.position = "right",
#         text=element_text(family="Times New Roman"),
#         plot.margin = unit(c(0,0,0,0), "cm"))
# a
# #obtain legend object
# legend = get_legend(a)
# a<-a+theme(legend.position = "right")
# ################ Risk table as ggplot element
# datatr <- fit.stan.as.ms10.f8$data %>%   ungroup() %>%
#   mutate(Strata = factor(
#     case_when(
#       RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
#       RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
#       RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
#       RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
#       .default = "other"
#     )
#   ))
# summary(datatr$Strata)
# patients<-datatr %>% 
#   group_by(PAT_MRN_ID) %>% 
#   arrange(tstart2) %>% 
#   slice_head()
# riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
#           RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
#           RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
#           RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
#     )
# 
# tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
#   label = as.character(value)
#   ))  +
#   geom_text(size = 4,family = "Times New Roman")+
#   coord_cartesian(xlim = c(0,60))+
#   scale_x_continuous(breaks=grid)+
#   scale_y_discrete(limits=rev(c(
#     "AA Comorbid",
#     "AA Non-Comorbid",
#     "EA Comorbid",
#     "EA Non-Comorbid"
#   )),labels=c("","","",""))+
#   labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
#   survminer::theme_survminer(base_family = "Times New Roman")+
#   theme(legend.position = "right",
#         text = element_text(family = "Times New Roman"),
#         axis.text.y = element_text( hjust = 1 ),
#         axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
#         axis.ticks.length.y = unit(15, "pt"),
#         plot.margin = unit(c(0,0,0,0), "cm"))
# 
# (a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```

##f9 plot

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test<-data_test %>% 
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
ps<-posterior_survfit(fit.stan.as.ms10.f9,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85,0.75),
  label=c("HR AA=1.16 (1.07 to 1.25)",
          "HR AA Comorbid=1.56 (1.48 to 1.64)",
          "HR EA Comorbid=1.42 (1.35 to 1.51)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "right")
################ Risk table as ggplot element
datatr <- fit.stan.as.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "right",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```







# Hazard for data_test

```{r}
(data_test<-data.frame(RACE = rep(c("AA","EA"),each=500),
                      age_visit=median(data$age_visit,na.rm=T),
                      dep_index=median(data$dep_index,na.rm=T),
                      avgDailyPM=median(data$avgDailyPM,na.rm=T),
                      ndvi=median(data$ndvi,na.rm=T),
                      onemore=rep(c(F,T),each=1000),
                      GENDER="Male",
                      time=rep(seq(0,60,length.out=500),4)
))

ndraws=1000
# already collapsed
ps1<-posterior_survfit(fit.stan.as.exp,newdata = data_test,extrapolate = F,times = "time",type = "haz",
                      draws = ndraws,return_matrix = F)
ps3<-posterior_survfit(fit.stan.as.ms10,newdata = data_test,extrapolate = F,times = "time",type = "haz",
                      draws = ndraws,return_matrix = F)

ps1<-ps1 %>% 
  mutate(RACE=rep(c("AA","EA","AA","EA"),each=500),
         onemore=rep(c("No","Yes"),each=1000))

ps3<-ps3 %>% 
  mutate(RACE=rep(c("AA","EA","AA","EA"),each=500),
         onemore=rep(c("No","Yes"),each=1000))

summary(factor(ps3$RACE))


a<-ps1 %>% as_tibble()%>%
  filter(RACE %in% c("AA","EA")) %>% 
  ggplot(aes(x=time,y=median,col=RACE,lty=onemore)) +
  geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub), fill = "gray90",size=0.1) +
  geom_line()+labs(x="Time (months)",y="Hazard rate",title="Constant (1000 draws)",lty="Comorbidities")+
  coord_cartesian(ylim = c(0,.09))+
  theme(legend.position = "none",plot.title = element_text(hjust = .5))+
  theme_minimal()

c<-ps3 %>% as_tibble()%>%
  filter(RACE %in% c("AA","EA")) %>% 
  ggplot(aes(x=time,y=median,col=RACE,lty=onemore))+
  geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub), fill = "gray90",size=0.1) +
  geom_line()+labs(x="Time (months)",y="",subtitle ="Hazard rate",lty="Comorbidities")+
  # coord_cartesian(ylim = c(0,.09))+
  theme(legend.position = "right",plot.title = element_text(hjust = .5))+
  theme_minimal()
c
a+c+plot_layout(guides = 'collect')
aa+c+plot_layout(guides = 'collect')


```


## RMST for M-Splines model

```{r}
data_test<-data.frame(id=1:2,
                      RACE = factor(c("AA","EA"),levels=c("EA","AA")),
                      hcType=factor(c("Medicaid","Medicaid"),levels=c("Commercial","Medicaid")),
                      age_visit=median(data$age_visit,na.rm = F),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=median(data$age_visit.cgm,na.rm = F),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=c(F,F),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )
tau <- c(12, 60)

rmst.as <-
  map(tau,
      ~ rmst_check_plot(
        fit.stan.as.ms10.f9,
        data_test,
        tau = .
      ))
ndig=1
rmst.table={}
for(i in 1:length(tau)) {
  aa=paste0(
    round(median(rmst.as[[i]][[1]]$rmstAA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$rmstAA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$rmstAA,prob=c(0.975)),ndig),
    ")"
  )
  ea=paste0(
    round(median(rmst.as[[i]][[1]]$rmstEA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$rmstEA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$rmstEA,prob=c(0.975)),ndig),
    ")"
  )
  diff=paste0(
    round(median(rmst.as[[i]][[1]]$diffAA_EA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$diffAA_EA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$diffAA_EA,prob=c(0.975)),ndig),
    ")"
  )
  obs=data.frame(tau=tau[i],
                   RMST.AA=aa,
                 RMST.EA=ea,
                 RMST.diff=diff
                 )
  rmst.table<-rbind(rmst.table,obs)
}
rmst.table%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
# join all 3 measures
rmst.as.gg<-rbind(
    rmst.as[[1]][[1]],
    rmst.as[[2]][[1]]#,
    # rmst.ar[[3]][[1]]
    )
rmst.as.gg$tau<-factor(as.character(rmst.as.gg$tau),levels=c("12","60"))
# wide to long for easy manipulation
rmst.as.gg<-gather(rmst.as.gg,condition,time,rmstAA:ratioAA_EA)

a<-ggplot() +
  geom_point(
    data = rmst.as.gg %>% filter(condition %in% c("rmstAA", "rmstEA")),
    aes(x = tau, y = time, group = condition,col=condition),
    position = "jitter",
    alpha = 0.05,
    size = 0.01
  ) +
  geom_boxplot()+
  stat_summary(
    # data=aa,
               # aes(x=tau,y=time,col=condition),
               fun=mean,
               geom="line"
               )+
  labs(y = "Time-free readmission",col="Group") + theme_bw()
a

bayesplot_grid(
    a,
    rmst.as[[1]]$p3,
    rmst.as[[2]]$p3,
    # rmst.ar[[3]]$p3,
    grid_args = list(ncol = 2),
    # titles = paste0("RMST (tau=", tau, ")"),
    # subtitles = rep("with medians and 95% CI", 4)
    subtitles = c("Time-free evolution","Tau=12","Tau=60")
    )
```




When comparing with comorbid.

```{r}
data_test<-data.frame(id=1:2,
                      RACE = factor(c("AA","EA"),levels=c("EA","AA")),
                      hcType=factor(c("Medicaid","Medicaid"),levels=c("Commercial","Medicaid")),
                      age_visit=median(data$age_visit,na.rm = F),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=median(data$age_visit.cgm,na.rm = F),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=c(T,T),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )

tau <- c(12, 60)

rmst.as <-
  map(tau,
      ~ rmst_check_plot(
        fit.stan.as.ms10.f9,
        data_test,
        tau = .
      ))
ndig=1
rmst.table={}
for(i in 1:length(tau)) {
  aa=paste0(
    round(median(rmst.as[[i]][[1]]$rmstAA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$rmstAA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$rmstAA,prob=c(0.975)),ndig),
    ")"
  )
  ea=paste0(
    round(median(rmst.as[[i]][[1]]$rmstEA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$rmstEA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$rmstEA,prob=c(0.975)),ndig),
    ")"
  )
  diff=paste0(
    round(median(rmst.as[[i]][[1]]$diffAA_EA),ndig),
    " (",
    round(quantile(rmst.as[[i]][[1]]$diffAA_EA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.as[[i]][[1]]$diffAA_EA,prob=c(0.975)),ndig),
    ")"
  )
  obs=data.frame(tau=tau[i],
                   RMST.AA=aa,
                 RMST.EA=ea,
                 RMST.diff=diff
                 )
  rmst.table<-rbind(rmst.table,obs)
}
rmst.table%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
# join all 3 measures
rmst.as.gg<-rbind(
    rmst.as[[1]][[1]],
    rmst.as[[2]][[1]]#,
    # rmst.ar[[3]][[1]]
    )
rmst.as.gg$tau<-factor(as.character(rmst.as.gg$tau),levels=c("12","60"))
# wide to long for easy manipulation
rmst.as.gg<-gather(rmst.as.gg,condition,time,rmstAA:ratioAA_EA)

a<-ggplot() +
  geom_point(
    data = rmst.as.gg %>% filter(condition %in% c("rmstAA", "rmstEA")),
    aes(x = tau, y = time, group = condition,col=condition),
    position = "jitter",
    alpha = 0.05,
    size = 0.01
  ) +
  geom_boxplot()+
  stat_summary(
    # data=aa,
               # aes(x=tau,y=time,col=condition),
               fun=mean,
               geom="line"
               )+
  labs(y = "Time-free readmission",col="Group") + theme_bw()
a

bayesplot_grid(
    a,
    rmst.as[[1]]$p3,
    rmst.as[[2]]$p3,
    # rmst.ar[[3]]$p3,
    grid_args = list(ncol = 2),
    # titles = paste0("RMST (tau=", tau, ")"),
    # subtitles = rep("with medians and 95% CI", 4)
    subtitles = c("Time-free evolution","Tau=12","Tau=60")
    )
```



## Validation



### C-Index

```{r}
ndraws=500
data_test<-fit.stan.as.exp.f4$data
data_test$coxlp.f4<-predict(cox.as4,newdata = data_test,"lp")
data_test$coxlp.f7<-predict(cox.as5,newdata = data_test,"lp")


#survival=exp(expected)
data_test$coxsurv.f4<-predict(cox.as4,newdata = data_test,"survival")
data_test$coxsurv.f7<-predict(cox.as5,newdata = data_test,"survival")

data_test$expsurv.f4<-posterior_survfit(fit.stan.as.exp.f4,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$expsurv.f7<-posterior_survfit(fit.stan.as.exp.f7,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$exploghaz.f4<-posterior_survfit(fit.stan.as.exp.f4,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$exploghaz.f7<-posterior_survfit(fit.stan.as.exp.f7,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$msloghaz10.f4<-posterior_survfit(fit.stan.as.ms10.f4,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$msloghaz10.f7<-posterior_survfit(fit.stan.as.ms10.f7,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$mssurv10.f4<-posterior_survfit(fit.stan.as.ms10.f4,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$mssurv10.f7<-posterior_survfit(fit.stan.as.ms10.f7,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
#f4
pairs(~coxlp.f4+exploghaz.f4+msloghaz10.f4, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)

pairs(~coxsurv.f4+expsurv.f4+mssurv10.f4, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)
#f5
pairs(~coxlp.f7+exploghaz.f7+msloghaz10.f7, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)

pairs(~coxsurv.f7+expsurv.f7+mssurv10.f7, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)
```

f4

```{r}
#data_test$mstime<-expre$time
y_test <- Surv(data_test$tstart,
               data_test$tstop,
               data_test$edvisit)

# cindex for linear predictor (log hazard)
concordance(y_test~data_test$coxlp.f4,reverse = T) #it works with risk
concordance(y_test~data_test$exploghaz.f4,data = data_test,reverse = T)
concordance(y_test~data_test$msloghaz10.f4,data = data_test,reverse = T)

```

f5

```{r}
#data_test$mstime<-expre$time
y_test <- Surv(data_test$tstart2,
               data_test$tstop2,
               data_test$edvisit)

# cindex for linear predictor (log hazard)
concordance(y_test~data_test$coxlp.f7,reverse = T) #it works with risk
concordance(y_test~data_test$exploghaz.f7,data = data_test,reverse = T)
concordance(y_test~data_test$msloghaz10.f7,data = data_test,reverse = T)
```


### Calibration plots

f4 

```{r}
#fixed time
times = as.double(seq(5, 60, 10))

summary(data_test$tstart)
#most common time slots
times2<-data_test %>% 
  dplyr::filter(tstart>0.0) %>% 
  mutate(ints = cut(tstart ,
                    breaks = seq(0, 60, 10),
                    include.lowest = FALSE,
                    right = FALSE)) %>% 
  dplyr::group_by(ints,tstart) %>% 
  dplyr::summarise(myn=n()) %>%  
  slice_max(myn, with_ties = F)
times2
times2<-times2%>% 
  pull(tstart) 
times
(times<-times2)

y_test.f4 <- filter(data_test, tstart %in% times) %>% 
  ungroup() %>% 
  select(c("tstart","tstop","edvisit")) 

```


```{r}
res<-calibrate(data = data_test,times = times,y = y_test.f4,
               tstart_col = "tstart",tstop_col ="tstop",status_col = "edvisit",
               n_groups = 10,surv_col = "mssurv10.f4" )

autoplot(res)
```

f5

```{r}
#fixed time
times = as.double(seq(5, 60, 10))

summary(data_test$tstart2)
#most common time slots
times2<-data_test %>% 
  dplyr::filter(tstart2>0.0) %>% 
  mutate(ints = cut(tstart2 ,
                    breaks = seq(0, 60, 10),
                    include.lowest = FALSE,
                    right = FALSE)) %>% 
  dplyr::group_by(ints,tstart2) %>% 
  dplyr::summarise(myn=n()) %>%  
  slice_max(myn, with_ties = F)
times2
times2<-times2%>% 
  pull(tstart2) 
times
(times<-times2)

y_test.f7 <- filter(data_test, tstart2 %in% times) %>% 
  ungroup() %>% 
  select(c("tstart2","tstop2","edvisit"))
```


```{r}
res<-calibrate(data = data_test,times = times,y = y_test.f7,
               tstart_col = "tstart2",tstop_col ="tstop2",status_col = "edvisit",
               n_groups = 10,surv_col = "mssurv10.f7" )

autoplot(res)
```




```{r}
# save.image("output/cchmc.as.v13.RData")

save(fit.stan.as.exp.f7,file="models/as.model.exp.f7.v13.RData")
save(fit.stan.as.ms10.f7,file="models/as.model.ms10.f7.v13.RData")
save(fit.stan.as.exp.f8,file="models/as.model.exp.f8.v13.RData")
save(fit.stan.as.ms10.f8,file="models/as.model.ms10.f8.v13.RData")
save(fit.stan.as.exp.f9,file="models/as.model.exp.f9.v13.RData")
save(fit.stan.as.ms10.f9,file="models/as.model.ms10.f9.v13.RData")
```
