---
title: "Untitled"
author: "Esteban Correa"
date: "2022-08-04"
output: html_document
---

v2 
bayes for asthma dataset

v3 
Make use of v5 dataset Generalization of rmst helper functions

v4 
Use onemore for one or more comordities AD/AS/AR/FA instead of Atopic dermatitis [Jocelyn's PARS paper].

v5
Updated to run on Apple Silicon 

v6
data/atm_survival_v7.RData with public insurance added

v7 
drop of deprivation index, EoE added to comorbidities

v8
cindex and calibration added
use of formula f4

v9
Use of atm_survival_v8.RData with improved inclusion criteria
code clean up, 10k MCMC, 95% CI for HR and RMST
use of formula f6

v10
use of formula f7

v11
Use of atm_survival_v8.RData 
use of f7 formula
Add f0 for null model
Code clean up and f4 models removed

v12
Re run for new splines2 v0.5.1 and rstanarm v2.26.1

v13
make use centered versions for greenspace and pm2.5
Inclusion of frailty term

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(rstanarm)
library(survival)
library(tidyverse)
library(patchwork)
library(bayesplot)
library(loo)
library(gtsummary)

source("helpers.r")

#To fix: Error in grid.Call(C_textBounds, as.graphicsAnnot(x$label), x$x, x$y,  : polygon edge not found
theme_set(theme_linedraw() +
            theme(text = element_text(family = "Times New Roman"),
                  panel.grid = element_blank()))

load("output/cchmc.ar.v13.RData")
```

# Asthma readmission

## Dataset preparation

```{r}
load("data/atm_survival_v8.RData")

# Note: imputation of covariates for last visit is required to enable right censoring
data<-ar_wide %>%
  mutate(GENDER=factor(GENDER,levels = c("Female", "Male")),
         RACE = factor(
    recode(
      RACE,
      `WHITE` = "EA",
      `BLACK OR AFRICAN AMERICAN` = "AA"
    ),
    levels = c("EA", "AA")
  )) %>%
  group_by(PAT_MRN_ID) %>%
  arrange(VISIT_DATE2) %>%
  mutate(
    lon = if_else(is.na(lon), lag(lon), lon),
    census_tract_id = if_else(
      is.na(census_tract_id),
      lag(census_tract_id),
      census_tract_id
    ),
    county_id = if_else(is.na(county_id), lag(county_id), county_id),
    dep_index = if_else(is.na(dep_index), lag(dep_index), dep_index),
    avgDailyPM = if_else(is.na(avgDailyPM), lag(avgDailyPM), avgDailyPM),
    ndvi = if_else(is.na(ndvi), lag(ndvi), ndvi),
    hcType = if_else(is.na(hcType), lag(hcType), hcType),
    comorbidities=rowSums(across(c(AD,AS,FA,EoE))), # except AR
    onemore=if_else(comorbidities>=1,TRUE,FALSE)
  )%>% 
  select(PAT_MRN_ID,tstart,tstop,tstart2,tstop2,edvisit,futime,enum,age_visit,GENDER,RACE,AD,FA,AS,AR,EoE,onemore,hcType,dep_index,ndvi,avgDailyPM) %>% 
  drop_na() %>% # added to avoid NA at validation
  arrange(PAT_MRN_ID,enum)

summary(ar_wide)
summary(data)

data2<-data %>% filter(tstart2==tstop2)
data<-data %>% 
  filter(!(PAT_MRN_ID %in% data2$PAT_MRN_ID))

data %>% 
  distinct(PAT_MRN_ID,.keep_all = T) %>% 
  group_by(RACE) %>% 
  summarise(freq=n())

#Centering
data <- data %>%
  ungroup() %>% 
  mutate(PAT_MRN_ID2=as.numeric(PAT_MRN_ID),
    age_visit.cgm = c(scale(age_visit, scale = T)),
    ndvi.cgm = c(scale(ndvi, scale = T)),
    avgDailyPM.cgm = c(scale(avgDailyPM, scale = T))
  ) %>%
  group_by(RACE) %>%
  mutate(ndvi.cwc = c(scale(ndvi, scale = T)),
         avgDailyPM.cwc = c(scale(avgDailyPM, scale = T))) %>%
  ungroup()
```

```{r}
# f0=formula(Surv(tstart, tstop, edvisit) ~ 1)
# cox.ar0<-coxph(formula = f0,data =data)
# # summary(cox.ar0)
# 
# f1=formula(Surv(tstart, tstop, edvisit) ~ RACE+AD+age_visit*GENDER+dep_index+ndvi+avgDailyPM)
# cox.ar<-coxph(formula = f1,data =data)
# # summary(cox.ar)
# 
# f2=formula(Surv(tstart, tstop, edvisit) ~ RACE+onemore+age_visit*GENDER+dep_index+ndvi+avgDailyPM)
# cox.ar2<-coxph(formula = f2,data =data)
# # summary(cox.ar2)
# 
# #public insurance
# f3=formula(Surv(tstart, tstop, edvisit) ~ RACE+onemore+hcType+age_visit*GENDER+dep_index+ndvi+avgDailyPM)
# cox.ar3<-coxph(formula = f3,data =data)
# # summary(cox.ar3)
# 
# #only public insurance
# f4=formula(Surv(tstart, tstop, edvisit) ~ age_visit*GENDER+RACE+hcType+onemore+ndvi+avgDailyPM)
# cox.ar4<-coxph(formula = f4,data =data)
# summary(cox.ar4)

######use of tstart2,tstop2
f0=formula(Surv(tstart2, tstop2, edvisit) ~ 1)
cox.ar0<-coxph(formula = f0,data =data)
# summary(cox.ar0)

f5=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit*GENDER+RACE+hcType+onemore+ndvi+avgDailyPM,type="interval")
cox.ar5<-coxph(formula = f5,data =data)
# summary(cox.ar5)
# use of interaction only
f6=formula(Surv(tstart2, tstop2, edvisit) ~ GENDER:age_visit+RACE+hcType+onemore+ndvi+avgDailyPM,type="interval")
cox.ar6<-coxph(formula = f6,data =data)
# summary(cox.ar6)

# use of interaction only for ndvi
f7=formula(Surv(tstart2, tstop2, edvisit) ~ age_visit:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi+RACE:avgDailyPM,type="interval")
cox.ar7<-coxph(formula = f7,data =data)
summary(cox.ar7)

# use of centered covariates
f8=formula(Surv(tstart2, tstop2, edvisit)~age_visit.cgm:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm,type="interval")
cox.ar8<-coxph(formula = f8,data =data)
summary(cox.ar8)

# use of centered covariates+frailty
f9=formula(Surv(tstart2, tstop2, edvisit) ~age_visit.cgm:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+frailty(PAT_MRN_ID),
           type="interval")
cox.ar9<-coxph(formula = f9,data =data)
summary(cox.ar9)

AIC(cox.ar0,cox.ar5,cox.ar6,cox.ar7,cox.ar8,cox.ar9)
cox.ar7%>%
  tbl_regression(exponentiate = TRUE)
cox.ar8%>%
  tbl_regression(exponentiate = TRUE)
cox.ar9%>%
  tbl_regression(exponentiate = TRUE)
```


## Stan version

```{r}
f8mcmc=formula(Surv(tstart2, tstop2, edvisit) ~age_visit.cgm:GENDER+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm,
           type="interval")

(f9mcmc=formula(Surv(tstart2, tstop2, edvisit) ~GENDER:age_visit.cgm+RACE+RACE:hcType+RACE:onemore+RACE:ndvi.cgm+RACE:avgDailyPM.cgm+(1|PAT_MRN_ID2),type="interval"))
# when chains>1 r makes use of viewer
CHAINS <- 4
CORES <- 4
ITER <- 1500
SEED <- 42
# draw from the prior predictive distribution of the stan_surv survival model
prior.stan.ar <- stan_surv(
  formula = f9mcmc,
  data = data ,
  basehaz = "exp",
  prior_PD = TRUE,
  chains = CHAINS,
  cores = CORES,
  iter = ITER,
  seed = SEED,
  refresh=1000,
  verbose=F)

# prior.stan.ar2 <- update(prior.stan.ar,
#                             prior_intercept = normal(0, 1),
#                             prior = normal(0, .2))
prior.stan.ar2 <- update(prior.stan.ar,
                            prior_intercept = normal(0, 5),
                            prior = normal(0, .2))

prior.stan.ar3 <- update(prior.stan.ar,
                            prior_intercept = cauchy(0,10),
                            prior = normal(0,.2))

prior.stan.ar4 <- update(prior.stan.ar,
                            prior_intercept = cauchy(0,25),
                            prior = normal(0,.2))
prior.stan.ar5 <- update(prior.stan.ar,
                            prior_intercept = cauchy(0,10),
                            prior = normal(0,.5))
prior.stan.ar6 <- update(prior.stan.ar,
                            prior_intercept = cauchy(0,20),
                            prior = normal(0,.5))
#more iterations
prior.stan.ar7 <- update(prior.stan.ar,
                         prior_intercept = normal(0,1),
                         prior = normal(0,.2),
                         iter = 15000)
prior.stan.ar8 <- update(prior.stan.ar,
                         prior_intercept = cauchy(0,10),
                         prior = normal(0,.2),
                         iter = 15000)
prior.stan.ar9 <- update(prior.stan.ar,
                         prior_intercept = cauchy(0,20),
                         prior = normal(0,.5),
                         iter = 15000)

prior.stan.ar10 <- update(prior.stan.ar,
                         prior_intercept = normal(0,2),
                         prior = student_t(3,0,1),
                         iter = 1500)

prior.stan.ar11 <- update(prior.stan.ar,
                         prior_intercept = student_t(3,0,1),
                         prior = student_t(3,0,1),
                         iter = 1500)
```

improving priors:

```{r}
# extract HR from classical coxph for arm=B
exp(coef(cox.ar9))[1]
base_cox_hr <- vline_at(exp(coef(cox.ar9))[1], color = "green")
a<-mcmc_hist(prior.stan.ar,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar1 default")
b<-mcmc_hist(prior.stan.ar2,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar2 N(0,1),N(0,.2))")
c<-mcmc_hist(prior.stan.ar3,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar3 C(10),N(0,.2))")
d<-mcmc_hist(prior.stan.ar4,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar4 C(25),N(0,.2))")
e<-mcmc_hist(prior.stan.ar5,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "Priors C(10),N(0,.5))")
f<-mcmc_hist(prior.stan.ar6,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar6 C(25),N(0,.5))")
g<-mcmc_hist(prior.stan.ar7,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar7 N(0,1),N(0,1),15k)")
h<-mcmc_hist(prior.stan.ar8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar8 N(0,1),N(0,.5),15k)")
i<-mcmc_hist(prior.stan.ar9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar9 C(25),N(0,.5),15k)")
j<-mcmc_hist(prior.stan.ar8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar10 st(3,0,1),N(0,.5),15k)")
k<-mcmc_hist(prior.stan.ar9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar11 N(0,1),st(3,0,1),15k)")
l<-mcmc_areas(prior.stan.ar10,
             pars = c("RACEAA"),
             transformations = exp,prob=0.95,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "ar11 N(0,1),st(3,0,1),15k)")
a+b+c+d+e+f+g+h+i+j+k+l+plot_layout(ncol = 3)
```

Prior ar9 selected:

```{r}
exp(coef(cox.ar9))

base_cox_hr<-vline_at(exp(coef(cox.ar9))[1], color = "green")
a<-mcmc_hist(prior.stan.ar6,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[2], color = "green")
b<-mcmc_hist(prior.stan.ar6,
             regex_pars = "Female",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[3], color = "green")
c<-mcmc_hist(prior.stan.ar6,
             regex_pars = "Male",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[4], color = "green")
d<-mcmc_hist(prior.stan.ar6,
             regex_pars = "RACEEA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[5], color = "green")
e<-mcmc_hist(prior.stan.ar6,
             regex_pars = "RACEAA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[6], color = "green")
f<-mcmc_hist(prior.stan.ar6,
             regex_pars = "RACEEA:onemore",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[7], color = "green")
g<-mcmc_dens(prior.stan.ar6,
             regex_pars = "RACEAA:onemore",
             transformations = exp,
             binwidth = 0.001,
             probs=c(.1,.5,.9)) +
  base_cox_hr+
  labs(subtitle = "Priors ar6")

a+b+c+d+e+f+g+plot_layout(ncol = 3)

```


Prior ar11 selected:

```{r}
exp(coef(cox.ar9))

base_cox_hr<-vline_at(exp(coef(cox.ar9))[1], color = "green")
a<-mcmc_hist(prior.stan.ar11,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[2], color = "green")
b<-mcmc_hist(prior.stan.ar11,
             regex_pars = "Female",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[3], color = "green")
c<-mcmc_hist(prior.stan.ar11,
             regex_pars = "Male",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[4], color = "green")
d<-mcmc_hist(prior.stan.ar11,
             regex_pars = "RACEEA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[5], color = "green")
e<-mcmc_hist(prior.stan.ar11,
             regex_pars = "RACEAA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[6], color = "green")
f<-mcmc_hist(prior.stan.ar11,
             regex_pars = "RACEEA:onemore",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[7], color = "green")
g<-mcmc_areas(prior.stan.ar11,
             regex_pars = "RACEAA:onemore",
             transformations = exp,
             binwidth = 0.001,
             prob=0.95) +
  base_cox_hr+
  labs(subtitle = "Priors ar11")
g
a+b+c+d+e+f+g+plot_layout(ncol = 3)

```


## Sampling exp+mspline

```{r}
f0
f8mcmc
f9mcmc
# exp
fit.stan.ar.exp.f7 <- update(prior.stan.ar6,  
                             prior_PD = FALSE,
                             formula=f7,
                             basehaz = "exp")

fit.stan.ar.exp.f8 <- update(prior.stan.ar6,  
                             prior_PD = FALSE,
                             formula=f8mcmc,
                             basehaz = "exp")
#based on ar2
fit.stan.ar.exp.f9 <- update(prior.stan.ar2,  
                             prior_PD = FALSE,
                             formula=f9mcmc,
                             basehaz = "exp")
#based on ar6 cauchy
fit.stan.ar.exp.f9b <- update(prior.stan.ar6,  
                             prior_PD = FALSE,
                             formula=f9mcmc,
                             chains = CHAINS,
                             cores = CORES,
                             iter = 3000,
                             basehaz = "exp")
#based on ar11 student-t
fit.stan.ar.exp.f9c <- update(prior.stan.ar11,  
                             prior_PD = FALSE,
                             formula=f9mcmc,
                             chains = CHAINS,
                             cores = CORES,
                             iter = 3000,
                             basehaz = "exp")

fit.stan.ar.exp.f0 <- update(prior.stan.ar6,  
                             prior_PD = FALSE,
                             data=fit.stan.ar.exp.f7$data,
                             formula=f0,
                             basehaz = "exp")

# cubic m-spline tstart2, tstop2
fit.stan.ar.ms10.f7 <- update(fit.stan.ar.exp.f7,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))

fit.stan.ar.ms10.f8 <- update(fit.stan.ar.exp.f8,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))

fit.stan.ar.ms10.f9 <- update(fit.stan.ar.exp.f9,
                        basehaz = "ms",
                        basehaz_ops = list(df = 10))

fit.stan.ar.ms10.f9b <- update(fit.stan.ar.exp.f9b,
                        basehaz = "ms",
                        chains = CHAINS,
                        cores = CORES,
                        iter = 3000,
                        basehaz_ops = list(df = 10))

fit.stan.ar.ms10.f9c <- update(fit.stan.ar.exp.f9c,
                        basehaz = "ms",
                        chains = CHAINS,
                        cores = CORES,
                        iter = 3000,
                        basehaz_ops = list(df = 10))

fits_stan <- list("exp.f0" = fit.stan.ar.exp.f0,
                  "exp.f7" = fit.stan.ar.exp.f7,
                  "ms10.f7" = fit.stan.ar.ms10.f7,
                  "exp.f8" = fit.stan.ar.exp.f8,
                  "ms10.f8" = fit.stan.ar.ms10.f8,
                  "exp.f9" = fit.stan.ar.exp.f9,
                  "ms10.f9" = fit.stan.ar.ms10.f9
                  )

#Are chains mixing well?
color_scheme_set("mix-blue-red")
mcmc_trace(fit.stan.ar.exp.f9, 
           pars=c("RACEAA", "(Intercept)"),
           facet_args = list(ncol = 1, strip.position = "left")
           )
```

## Posterior uncertainty intervals


```{r}
mcmc_post_ci(fit.stan.ar.exp.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.exp.f9b,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.exp.f9c,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
```


```{r}
mcmc_post_ci(fit.stan.ar.ms10.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.ms10.f9b,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.ms10.f9c,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
```



```{r}
mcmc_post_ci(fit.stan.ar.ms10.f7,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.ms10.f8,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
mcmc_post_ci(fit.stan.ar.ms10.f9,.95,12)%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 

```

## Hazard ratio checking

```{r}
# extract HR from classical coxph for arm=B
exp(coef(cox.ar9)[1])
base_cox_hr <- vline_at(exp(coef(cox.ar9))[1], color = "green")


a<-mcmc_hist(prior.stan.ar6,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle="f0")

b<-mcmc_hist(fit.stan.ar.exp.f7,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f7 (exp)")
c<-mcmc_hist(fit.stan.ar.ms10.f7,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f7 (ms-10)")

d<-mcmc_hist(fit.stan.ar.exp.f8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f8 (exp)")
e<-mcmc_hist(fit.stan.ar.ms10.f8,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f8 (ms-10)")
f<-mcmc_hist(fit.stan.ar.exp.f9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f9 (exp)")
g<-mcmc_hist(fit.stan.ar.ms10.f9,
             pars = c("RACEAA"),
             transformations = exp,
             binwidth = 0.001) + base_cox_hr+labs(subtitle = "f9 (ms-10)")

a+plot_spacer()+b+c+d+e+f+g+plot_layout(ncol = 4)
```



```{r}
exp(coef(cox.ar9))

base_cox_hr<-vline_at(exp(coef(cox.ar9))[1], color = "green")
a<-mcmc_areas(fit.stan.ar.ms10.f9c,
             pars = c("RACEAA"),
             transformations = exp,prob=0.95,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[2], color = "green")
b<-mcmc_hist(fit.stan.ar.ms10.f9c, 
             regex_pars = "Female",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[3], color = "green")
c<-mcmc_hist(fit.stan.ar.ms10.f9c,
             regex_pars = "Male",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[4], color = "green")
d<-mcmc_hist(fit.stan.ar.ms10.f9c,
             regex_pars = "RACEEA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[5], color = "green")
e<-mcmc_hist(fit.stan.ar.ms10.f9c,
             regex_pars = "RACEAA:hcTypeMedicaid",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")

base_cox_hr<-vline_at(exp(coef(cox.ar9))[6], color = "green")
f<-mcmc_hist(fit.stan.ar.ms10.f9c,
             regex_pars = "RACEEA:onemore",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")
base_cox_hr<-vline_at(exp(coef(cox.ar9))[7], color = "green")
g<-mcmc_hist(fit.stan.ar.ms10.f9c,
             regex_pars = "RACEAA:onemore",
             transformations = exp,
             binwidth = 0.001) +
  base_cox_hr+
  labs(subtitle = "f9c")

a+b+c+d+e+f+g+plot_layout(ncol = 3)

```



## Hazard baseline

```{r}
plots <- map(fits_stan,plot)

a<-plots[[1]]+
  labs(title = "f0 (exp)")+
  coord_cartesian(ylim = c(0,.1))+
  theme(plot.title = element_text(hjust = .5))

b<-plots[[2]]+labs(title = "f7  (exp)")+
  coord_cartesian(ylim = c(0,.1))+
  theme(plot.title = element_text(hjust = .5))

c<-plots[[3]]+labs(title = "f7  (ms-10)")+
  coord_cartesian(ylim = c(0,.9))+
  theme(plot.title = element_text(hjust = .5))

d<-plots[[4]]+labs(title = "f8 (exp)")+
  coord_cartesian(ylim = c(0,.1))+
  theme(plot.title = element_text(hjust = .5))

e<-plots[[5]]+labs(title = "f8 (ms-10)")+
  coord_cartesian(ylim = c(0,.9))+
  theme(plot.title = element_text(hjust = .5))

f<-plots[[6]]+labs(title = "f9 (exp)")+
  coord_cartesian(ylim = c(0,.1))+
  theme(plot.title = element_text(hjust = .5))

g<-plots[[7]]+labs(title = "f9 (ms-10)")+
  coord_cartesian(ylim = c(0,.9))+
  theme(plot.title = element_text(hjust = .5))

a+b+c+d+e+f+g

```

# Survival curves: Posterior vs COX PH

```{r}
# f9 centered+frailty
data_test<-data.frame(id=1:4,
                      RACE = factor(rep(c("AA","EA"),each=2),levels=c("EA","AA")),
                      hcType=factor("Medicaid",levels=c("Commercial","Medicaid")),
                      age_visit=rep(median(data$age_visit,na.rm = F),2),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=rep(median(data$age_visit.cgm,na.rm = F),2),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=rep(c(F,T),2),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )
```

##f7 plot

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test<-data_test %>% 
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
ps<-posterior_survfit(fit.stan.ar.ms10.f7,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA Comorbid=1.94 (1.86 to 2.02)",
          "HR EA Comorbid=2.15 (2.07 to 2.23)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "right")
################ Risk table as ggplot element
datatr <- fit.stan.ar.ms10.f7$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "right",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```


##f8 plot

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test<-data_test %>% 
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
ps<-posterior_survfit(fit.stan.ar.ms10.f8,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c("HR AA Comorbid=1.94 (1.85 to 2.02)",
          "HR EA Comorbid=2.15 (2.07 to 2.23)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "right")
################ Risk table as ggplot element
datatr <- fit.stan.ar.ms10.f8$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "right",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```

##f9 plot

```{r}
annoTextS=4
cbPalette <- c("#E64B35","#d47911", "#4DBBD5", "#009E73")
grid <- seq(0,60,by=12)

data_test<-data_test %>% 
  mutate(Strata = paste0(RACE, ifelse(onemore, " Comorbid", " Non-comorbid")))
ndraws=500
# already collapsed
ps<-posterior_survfit(fit.stan.ar.ms10.f9,  
                      newdata = data_test,
                      times = 0,
                      extrapolate   = T, 
                      condition     = FALSE,
                      return_matrix = F,
                      control = list(edist = 60),
                      draws = ndraws)
ps<-ps %>% 
  left_join(data_test,by="id")
# prepare HR annotations
text.df<-data.frame(
  x=c(15),
  y=c(0.95,0.85),
  label=c(
          "HR AA Comorbid=1.62 (1.52 to 1.72)",
          "HR EA Comorbid=1.55 (1.46 to 1.64)")
)
################ survival curves
a<-ps %>%as_tibble()%>%
  ggplot(aes(x=time,y=median,col=Strata)) +
  # geom_ribbon(aes(ymin = ci_lb, ymax = ci_ub,fill=Strata),
  #             # fill = "gray90",
  #             alpha=0.2,
  #             size=0.0) +
  geom_line()+
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values = cbPalette)+
  scale_x_continuous(breaks = grid)+
  labs(x="",y="Probability of Survival Free\n of ED Encounter",subtitle="A",col="Strata")+
  survminer::theme_survminer(base_family = "Times New Roman")
a<-a+annotate(geom="text",x=text.df$x,y=text.df$y,label=text.df$label,
           size=annoTextS,
           hjust=0,family="Times New Roman")+
  theme(legend.position = "right",
        text=element_text(family="Times New Roman"),
        plot.margin = unit(c(0,0,0,0), "cm"))
a
#obtain legend object
legend = get_legend(a)
a<-a+theme(legend.position = "right")
################ Risk table as ggplot element
datatr <- fit.stan.ar.ms10.f9$data %>%   ungroup() %>%
  mutate(Strata = factor(
    case_when(
      RACE == "AA" & onemore == TRUE ~ "AA Comorbid",
      RACE == "EA" & onemore == TRUE ~ "EA Comorbid",
      RACE == "AA" & onemore == F ~ "AA Non-Comorbid",
      RACE == "EA" & onemore == F ~ "EA Non-Comorbid",
      .default = "other"
    )
  ))
summary(datatr$Strata)
patients<-datatr %>% 
  group_by(PAT_MRN_ID) %>% 
  arrange(tstart2) %>% 
  slice_head()
riskcounts.df<-rbind(RiskSetCount(grid,patients,strataoi ="AA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="AA Non-Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Comorbid"),
          RiskSetCount(grid,patients,strataoi ="EA Non-Comorbid")
    )

tabrisk<-ggplot(riskcounts.df, aes(x = time,y = factor(strata),
  label = as.character(value)
  ))  +
  geom_text(size = 4,family = "Times New Roman")+
  coord_cartesian(xlim = c(0,60))+
  scale_x_continuous(breaks=grid)+
  scale_y_discrete(limits=rev(c(
    "AA Comorbid",
    "AA Non-Comorbid",
    "EA Comorbid",
    "EA Non-Comorbid"
  )),labels=c("","","",""))+
  labs(x="Time (months)",y="Strata",subtitle = "Number at risk")+
  survminer::theme_survminer(base_family = "Times New Roman")+
  theme(legend.position = "right",
        text = element_text(family = "Times New Roman"),
        axis.text.y = element_text( hjust = 1 ),
        axis.ticks.y = element_line(size  = 2,colour = rev(cbPalette)),
        axis.ticks.length.y = unit(15, "pt"),
        plot.margin = unit(c(0,0,0,0), "cm"))

(a<-a / tabrisk+plot_layout(ncol=1,heights = c(3,1)))
```



## RMST for M-Splines model

When conditioning on race

```{r}
data_test<-data.frame(id=1:2,
                      RACE = factor(c("AA","EA")),
                      hcType=factor(c("Medicaid","Medicaid"),levels=c("Commercial","Medicaid")),
                      age_visit=median(data$age_visit,na.rm = F),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=median(data$age_visit.cgm,na.rm = F),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=c(F,F),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )
tau <- c(12, 60)

rmst.ar <-
  map(tau,
      ~ rmst_check_plot(
        fit.stan.ar.ms10.f9,
        data_test,
        tau = .
      ))
ndig=1
rmst.table={}
for(i in 1:length(tau)) {
  aa=paste0(
    round(median(rmst.ar[[i]][[1]]$rmstAA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$rmstAA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$rmstAA,prob=c(0.975)),ndig),
    ")"
  )
  ea=paste0(
    round(median(rmst.ar[[i]][[1]]$rmstEA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$rmstEA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$rmstEA,prob=c(0.975)),ndig),
    ")"
  )
  diff=paste0(
    round(median(rmst.ar[[i]][[1]]$diffAA_EA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$diffAA_EA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$diffAA_EA,prob=c(0.975)),ndig),
    ")"
  )
  obs=data.frame(tau=tau[i],
                   RMST.AA=aa,
                 RMST.EA=ea,
                 RMST.diff=diff
                 )
  rmst.table<-rbind(rmst.table,obs)
}
rmst.table%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
# join all 3 measures
rmst.ar.gg<-rbind(
    rmst.ar[[1]][[1]],
    rmst.ar[[2]][[1]]#,
    # rmst.ar[[3]][[1]]
    )
rmst.ar.gg$tau<-factor(as.character(rmst.ar.gg$tau),levels=c("12","60"))
# wide to long for easy manipulation
rmst.ar.gg<-gather(rmst.ar.gg,condition,time,rmstAA:ratioAA_EA)

a<-ggplot() +
  geom_point(
    data = rmst.ar.gg %>% filter(condition %in% c("rmstAA", "rmstEA")),
    aes(x = tau, y = time, group = condition,col=condition),
    position = "jitter",
    alpha = 0.05,
    size = 0.01
  ) +
  geom_boxplot()+
  stat_summary(
    # data=aa,
               # aes(x=tau,y=time,col=condition),
               fun=mean,
               geom="line"
               )+
  labs(y = "Time-free readmission",col="Group") + theme_bw()
a

bayesplot_grid(
    a,
    rmst.ar[[1]]$p3,
    rmst.ar[[2]]$p3,
    # rmst.ar[[3]]$p3,
    grid_args = list(ncol = 2),
    # titles = paste0("RMST (tau=", tau, ")"),
    # subtitles = rep("with medians and 95% CI", 4)
    subtitles = c("Time-free evolution","Tau=12","Tau=60")
    )

```


When comparing on race with comorbidities.

```{r}
data_test<-data.frame(id=1:2,
                      RACE = factor(c("AA","EA")),
                      hcType=factor(c("Medicaid","Medicaid"),levels=c("Commercial","Medicaid")),
                      age_visit=median(data$age_visit,na.rm = F),
                      avgDailyPM=median(data$avgDailyPM,na.rm = F),
                      ndvi=median(data$ndvi,na.rm = F),
                      age_visit.cgm=median(data$age_visit.cgm,na.rm = F),
                      avgDailyPM.cgm=median(data$avgDailyPM.cgm,na.rm = F),
                      ndvi.cgm=median(data$ndvi.cgm,na.rm = F),
                      onemore=c(T,T),
                      GENDER=factor(rep("Male",2),levels=c("Female","Male")),
                      PAT_MRN_ID2=as.numeric(1)
                      )
tau <- c(12, 60)

rmst.ar <-
  map(tau,
      ~ rmst_check_plot(
        fit.stan.ar.ms10.f9,
        data_test,
        tau = .
      ))
ndig=1
rmst.table={}
for(i in 1:length(tau)) {
  aa=paste0(
    round(median(rmst.ar[[i]][[1]]$rmstAA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$rmstAA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$rmstAA,prob=c(0.975)),ndig),
    ")"
  )
  ea=paste0(
    round(median(rmst.ar[[i]][[1]]$rmstEA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$rmstEA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$rmstEA,prob=c(0.975)),ndig),
    ")"
  )
  diff=paste0(
    round(median(rmst.ar[[i]][[1]]$diffAA_EA),ndig),
    " (",
    round(quantile(rmst.ar[[i]][[1]]$diffAA_EA,prob=c(0.025)),ndig),
    " to ",
    round(quantile(rmst.ar[[i]][[1]]$diffAA_EA,prob=c(0.975)),ndig),
    ")"
  )
  obs=data.frame(tau=tau[i],
                   RMST.AA=aa,
                 RMST.EA=ea,
                 RMST.diff=diff
                 )
  rmst.table<-rbind(rmst.table,obs)
}
rmst.table%>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover",full_width=F) 
# join all 3 measures
rmst.ar.gg<-rbind(
    rmst.ar[[1]][[1]],
    rmst.ar[[2]][[1]]#,
    # rmst.ar[[3]][[1]]
    )
rmst.ar.gg$tau<-factor(as.character(rmst.ar.gg$tau),levels=c("12","60"))
# wide to long for easy manipulation
rmst.ar.gg<-gather(rmst.ar.gg,condition,time,rmstAA:ratioAA_EA)

a<-ggplot() +
  geom_point(
    data = rmst.ar.gg %>% filter(condition %in% c("rmstAA", "rmstEA")),
    aes(x = tau, y = time, group = condition,col=condition),
    position = "jitter",
    alpha = 0.05,
    size = 0.01
  ) +
  geom_boxplot()+
  stat_summary(
    # data=aa,
               # aes(x=tau,y=time,col=condition),
               fun=mean,
               geom="line"
               )+
  labs(y = "Time-free readmission",col="Group") + theme_bw()
a

bayesplot_grid(
    a,
    rmst.ar[[1]]$p3,
    rmst.ar[[2]]$p3,
    # rmst.ar[[3]]$p3,
    grid_args = list(ncol = 2),
    # titles = paste0("RMST (tau=", tau, ")"),
    # subtitles = rep("with medians and 95% CI", 4)
    subtitles = c("Time-free evolution","Tau=12","Tau=60")
    )

```




## Validation

### LOGO ELPD

```{r}
post <- as.array(fit.stan.ar.ms10) 
ids<-fit.stan.ar.exp$data$PAT_MRN_ID
chain_id=rep(1:dim(post)[2], each = dim(post)[1])
#model exp
myllk_exp<-log_lik(fit.stan.ar.exp,merge_chains = F)
# join llk by patient
myllk2_exp<-llkByPatient(llk = myllk_exp,ids = ids)
reff_exp<-relative_eff(myllk2_exp,chain_id = chain_id,cores = 4)
#model ms10
myllk_ms10<-log_lik(fit.stan.ar.ms10)
myllk2_ms10<-llkByPatient(llk = myllk_ms10,ids = ids)
reff_ms10<-relative_eff(myllk2_ms10,chain_id = chain_id,cores = 4)
# #model ms10b
myllk_ms10b<-log_lik(fit.stan.ar.ms10)
myllk2_ms60<-llkByPatient(llk = myllk_ms60,ids = ids)
reff_ms60<-relative_eff(myllk2_ms60,chain_id = chain_id,cores = 4)

compare(loo(myllk2_exp,r_eff = reff_exp),
            loo(myllk2_ms10,r_eff = reff_ms10),
            loo(myllk2_ms60,r_eff = reff_ms60)
            )
compare(waic(myllk2_exp,r_eff = reff_exp),
            waic(myllk2_ms10,r_eff = reff_ms10),
            waic(myllk2_ms60,r_eff = reff_ms60)
            )
```

### C-Index

```{r}
ndraws=500
data_test<-fit.stan.ar.exp.f8$data

data_test$coxlp.f7<-predict(cox.ar7,newdata = fit.stan.ar.exp.f7$data,"lp")
data_test$coxsurv.f7<-predict(cox.ar7,newdata = fit.stan.ar.exp.f7$data,"survival")
data_test$coxlp.f8<-predict(cox.ar8,newdata = data_test,"lp")
data_test$coxsurv.f8<-predict(cox.ar8,newdata = data_test,"survival")
# data_test$coxlp.f9<-predict(cox.ar9,newdata = fit.stan.ar.exp.f9$data,"lp")
# data_test$coxsurv.f9<-predict(cox.ar9,newdata = data_test,"survival")

#f7
data_test$exploghaz.f7<-posterior_survfit(fit.stan.ar.exp.f7,
                      newdata = fit.stan.ar.exp.f7$data,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$expsurv.f7<-posterior_survfit(fit.stan.ar.exp.f7,
                      newdata = fit.stan.ar.exp.f7$data,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$msloghaz10.f7<-posterior_survfit(fit.stan.ar.ms10.f7,
                      newdata = fit.stan.ar.exp.f7$data,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$mssurv10.f7<-posterior_survfit(fit.stan.ar.ms10.f7,
                      newdata = fit.stan.ar.exp.f7$data,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

#f8
data_test$exploghaz.f8<-posterior_survfit(fit.stan.ar.exp.f8,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$expsurv.f8<-posterior_survfit(fit.stan.ar.exp.f8,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$msloghaz10.f8<-posterior_survfit(fit.stan.ar.ms10.f8,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$mssurv10.f8<-posterior_survfit(fit.stan.ar.ms10.f8,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
#f9
data_test$exploghaz.f9<-posterior_survfit(fit.stan.ar.exp.f9,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$expsurv.f9<-posterior_survfit(fit.stan.ar.exp.f9,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

data_test$msloghaz10.f9<-posterior_survfit(fit.stan.ar.ms10.f9,
                      newdata = data_test,
                      extrapolate = F,
                      type="loghaz",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median
data_test$mssurv10.f9<-posterior_survfit(fit.stan.ar.ms10.f9,
                      newdata = data_test,
                      extrapolate = F,
                      type="surv",
                      draws = ndraws,return_matrix = F,
                      times       = "tstart",
                      last_time   = "tstop")$median

pairs(~coxlp.f8+coxlp.f9+exploghaz.f8+msloghaz10.f8+exploghaz.f9+msloghaz10.f9, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)

pairs(~coxsurv.f8+coxsurv.f9+expsurv.f8+mssurv10.f8+expsurv.f9+mssurv10.f9, data_test,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth)
```

```{r}
#data_test$mstime<-expre$time
y_test <- Surv(data_test$tstart2,
               data_test$tstop2,
               data_test$edvisit)

# cindex for linear predictor (log hazard)
concordance(y_test~data_test$coxlp.f7,reverse = T) #it works with risk
concordance(y_test~data_test$exploghaz.f7,data = data_test,reverse = T)
concordance(y_test~data_test$coxlp.f8,reverse = T) #it works with risk
concordance(y_test~data_test$exploghaz.f8,data = data_test,reverse = T)
concordance(y_test~data_test$msloghaz10.f8,data = data_test,reverse = T)
concordance(y_test~data_test$exploghaz.f9,data = data_test,reverse = T)
concordance(y_test~data_test$msloghaz10.f9,data = data_test,reverse = T)

```




### Calibration plots

```{r}
#fixed time
times = as.double(seq(5, 60, 10))

summary(data_test$tstart2)
#most common time slots
times2<-data_test %>% 
  dplyr::filter(tstart2>0.0) %>% 
  mutate(ints = cut(tstart2,
                    breaks = seq(0, 60, 10),
                    include.lowest = FALSE,
                    right = FALSE)) %>% 
  dplyr::group_by(ints,tstart2) %>% 
  dplyr::summarise(myn=n()) %>%  
  slice_max(myn, with_ties = F)
times2
times2<-times2%>% 
  pull(tstart2) 
times
(times<-times2)

y_test1 <- filter(data_test, tstart2 %in% times) %>% 
  ungroup() %>% 
  select(c("tstart2","tstop2","edvisit")) 

```


```{r}
library(data.table)
res<-calibrate(data = data_test,times = times,y = y_test1,
               n_groups = 10,surv_col = "mssurv10" )

autoplot(res)
```


```{r}
# save.image("output/cchmc.ar.v13.RData")

save(fit.stan.ar.exp.f7,file="models/ar.model.exp.f7.v13.RData")
save(fit.stan.ar.ms10.f7,file="models/ar.model.ms10.f7.v13.RData")
save(fit.stan.ar.exp.f8,file="models/ar.model.exp.f8.v13.RData")
save(fit.stan.ar.ms10.f8,file="models/ar.model.ms10.f8.v13.RData")
save(fit.stan.ar.exp.f9,file="models/ar.model.exp.f9.v13.RData")
save(fit.stan.ar.ms10.f9,file="models/ar.model.ms10.f9.v13.RData")
```


